---
title: "Chap6_05_24_2019"
author: "Kazu"
date: "5/16/2019"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=TRUE)
#install.packages(c("devtools","mvtnorm","loo","coda"),dependencies=TRUE)
#library(devtools)
#install_github("rmcelreath/rethinking",ref="Experimental")
library(rethinking)
library(lme4);library(lmerTest);library(ggplot2);library(reshape2);library(tidyverse);library(readr)
```

Overthinking: Simulated science distortion. Simulations like this one are easy to do in R, or in any other scripting language, once you have seen a few examples. In this simulation, we just draw some random Gaussian criteria for a sample of proposals and then select the top 10% combined scores.
# R code 6.1
```{r}
set.seed(1914)
N <- 200 # num grant proposals
p <- 0.1 # proportion to select
# uncorrelated newsworthiness and trustworthiness 
nw <- rnorm(N)
tw <- rnorm(N)
# select top 10% of combined scores
s<-nw+tw #totalscore
q <- quantile( s , 1-p ) # top 10% threshold 
selected <- ifelse( s >= q , TRUE , FALSE )
cor( tw[selected] , nw[selected] )
```

# 6.1. Multicollinearity
6.1.1. Multicollinear legs.
The code below will simulate the heights and leg lengths of 100 individuals.
```{r}
N <- 100  # number of individuals
set.seed(909)
height <- rnorm(N,10,2)
leg_prop <- runif(N,0.4,0.5)

```



