---
title: "Chap14_5_July7_2020"
author: "Kazu"
date: "7/5/2020"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=TRUE,cache = FALSE)
```

# 1) Load the data as follows:
```{r}
library(tidyverse)
library(rethinking)
library(ape)
library(ade4)
data("carni70")
tre <- read.tree(text=carni70$tre)
d <- carni70$tab
?carni70
# "This data set describes the phylogeny of 70 carnivora as reported by Diniz-Filho and Torres (2002). It also gives the geographic range size and body size corresponding to these 70 species."
```

# 2) Note that the species name given in "rownames(d)" and in "tre$tip.label" do not match.  Fix that (e.g. using str_replace() )
```{r}
rownames(d) # "_"
tre$tip.label # "."  
tre$tip.label %in% str_replace(rownames(d),"_","\\.") # all "TRUE"
# copied from R code 14.47
spp_obs <- str_replace(rownames(d), pattern="_", replacement = ".")
tree_trimmed <- keep.tip( tre, spp_obs )

```

# 3) Should either range or size be transformed?
```{r}
d %>% ggplot(aes(x=size)) + geom_histogram()
d %>% ggplot(aes(x=range)) + geom_histogram()
# When I need to think about tranformation?

# correlation between size and range?
d %>% ggplot(aes(x=log(size),y=log(range))) + geom_point()
             
# I do not know. Anyway standardize? (see R code 14.48)
dat_list <- list(
    N_spp = nrow(d),
    logsize = standardize(log(d$size)),
    logrange = standardize(log(d$range)),
    Imat = diag( nrow(d) )
)

# R code 14.49
Rbm <- corBrownian( phy=tree_trimmed )
V <- vcv(Rbm) # warning
Dmat <- cophenetic( tree_trimmed )
plot( Dmat , V , xlab="phylogenetic distance" , ylab="covariance" )
```

# 4) Does size predict range?  Evaluate this question without and with accounting for phylogenetic relationships.

# copy and modified from R code 14.50
```{r}
# put species in right order
dat_list$V <- V[ spp_obs , spp_obs ]
# convert to correlation matrix
dat_list$R <- dat_list$V / max(V)
# add scaled and reordered distance matrix
dat_list$Dmat <- Dmat[ spp_obs , spp_obs ] / max(Dmat)

# Brownian motion model
m14.9.mod <- ulam(
    alist(
        logrange ~ multi_normal( mu , SIGMA ),
        mu <- a + b*logsize,
        matrix[N_spp,N_spp]: SIGMA <- R * sigma_sq,
        a ~ normal( 0 , 1 ),
        b ~ normal( 0 , 0.5 ),
        sigma_sq ~ exponential( 1 )
    ), data=dat_list , chains=4 , cores=4)
precis( m14.9.mod )
```

# Copy and modified R coe 14.51
```{r R code 14.51}
# add scaled and reordered distance matrix
dat_list$Dmat <- Dmat[ spp_obs , spp_obs ] / max(Dmat)
m14.10.mod <- ulam(
    alist(
        logrange ~ multi_normal( mu , SIGMA ),
        mu <- a + b*logsize,
        matrix[N_spp,N_spp]: SIGMA <- cov_GPL2( Dmat , etasq , rhosq , 0.01 ),
        a ~ normal(0,1),
        b ~ normal(0,0.5),
        etasq ~ exponential(1),
        rhosq ~ exponential(1)
    ), data=dat_list , chains=4 , cores=4 )
precis( m14.10.mod )
```
* Since b is positive and slighly higher (0.39 or 0.36) in both models, log size matters.

# 5) Discuss your findings.
```{r}

```

