---
title: "Chap13_Jan31_2020_HW_Kazu"
author: "Kazu"
date: "1/31/2020"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

# 13. Models with memory
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=TRUE,cache = FALSE)
library(rethinking)
library(lme4);library(lmerTest);library(ggplot2);library(reshape2);library(tidyverse);library(readr)
```

Please find attached csv.  This csv has measurements of tomato internodes and petioles from wildtype (Moneymaker) and various phytochrome mutant lines.  Measurements were made at 3 time points, 21, 28, and 35 days after germination under two treatments, simulated sun and simulated shade.

for today let's focus on day 35.  Also let's focus on total stem length.  So: first

# Q1)
## a) subset the data for day 35
```{r}
F4phyE<-read_csv("../figure4phyE.csv")
summary(F4phyE)
F4phyE.day35 <- F4phyE %>% filter(day==35)
summary(F4phyE.day35)
```

## b) create a new column "stem_length" that is the sum of epi, int1, int2, and int3
```{r}
F4phyE.day35.v2 <- F4phyE.day35 %>% mutate(stem_length=epi+int1+int2+int3) # %>% View()
F4phyE.day35.v2
```

c) although flats are listed as 1-6, flats in sun and shade are separate. Create a new column "flat2" that corrects for this.
```{r}
F4phyE.day35.v3 <- F4phyE.day35.v2 %>% unite("flat2",c("treatment","flat"),sep="_", remove=FALSE)
F4phyE.day35.v3
```


Ultimately you want to know if any of the mutants have a different length from Moneymaker, in sun or in shade, or if the response to shade differs.
```{r}
dat_list.Q1error<- list(
    stem_length = F4phyE.day35.v3$stem_length,
    genotype = as.integer(factor(F4phyE.day35.v3$genotype,levels=c("Moneymaker", 
                                            "phyB1", 
                                            "phyB2", 
                                            "phyB1/B2", 
                                            "phyEami3", 
                                            "phyEami7"))),
    flat_id = as.integer(factor(F4phyE.day35.v3$flat2,levels=c("shade_1",
                                                            "shade_2",
                                                            "shade_3",
                                                            "shade_4",
                                                            "shade_5",
                                                            "shade_6",
                                                            "sun_1",
                                                            "sun_2",
                                                            "sun_3",
                                                            "sun_4",
                                                            "sun_5",
                                                            "sun_6"))),
    treatment = ifelse(F4phyE.day35.v3$treatment=="sun",0,1)) 
dat_list.Q1<- list(
    stem_length = F4phyE.day35.v3$stem_length,
    genotype = as.integer(factor(F4phyE.day35.v3$genotype,levels=c("Moneymaker", 
                                            "phyB1", 
                                            "phyB2", 
                                            "phyB1/B2", 
                                            "phyEami3", 
                                            "phyEami7"))),
    flat_id = as.integer(factor(F4phyE.day35.v3$flat2,levels=c("shade_1",
                                                            "shade_2",
                                                            "shade_3",
                                                            "shade_4",
                                                            "shade_5",
                                                            "shade_6",
                                                            "sun_1",
                                                            "sun_2",
                                                            "sun_3",
                                                            "sun_4",
                                                            "sun_5",
                                                            "sun_6"))),
    treatment = ifelse(F4phyE.day35.v3$treatment=="sun",1,2)) 
# "sun" is 1, "shade" is 2
dat_list.Q1
str(dat_list.Q1error)
str(dat_list.Q1)

```

# look the raw data
```{r}
str(F4phyE.day35.v3)
F4phyE.day35.v3$treatment <-factor(F4phyE.day35.v3$treatment,levels=c("sun","shade"))
F4phyE.day35.v3 %>% View()
F4phyE.day35.v3 %>% ggplot(aes(x=treatment, y=stem_length,color=treatment)) + geom_jitter() + facet_grid(.~genotype)
```
* Looks there are treatment effects are genotype effects.

# Q2) Fit 3 models, all of which include genotype and treatment
a) do not include flat
## ModelQ1zero
$$
stem_length_i \sim Normal(mu,sigma) \\
mu = \theta + \alpha_{genotype} + \beta_{treatment} \\
Kazu'Q: When\ logit\ should\ be\ used? \\
\theta \sim Normal(0,1) \\
\alpha_{group} \sim Normal(0, 1) \\
\beta_{treatment} \sim Normal(0, 1)
$$

## ModelQ1a
$$
stem_length_i \sim Normal(0,p) \\
logit(p) = \alpha_{genotype} + \beta_{treatment} \\
Kazu'Q: When\ logit\ should\ be\ used? \\

\alpha_{group} \sim Normal(0, 1) \\
\beta_{treatment} \sim Normal(0, 1)
$$


```{r}
set.seed(15)
# error
mQ1.a.error <- ulam(
    alist(
        stem_length ~ dnorm( 1 , p),
        logit(p) <- a[genotype] + b[treatment],
        b[treatment] ~ dnorm( 0 , 1),
        a[genotype] ~ dnorm(0 , 1)
    ), data=dat_list.Q1error , chains=2 , cores=2 , log_lik=TRUE )
# no error, why?
mQ1.a <- ulam(
    alist(
        stem_length ~ dnorm( 1 , p),
        logit(p) <- a[genotype] + b[treatment],
        b[treatment] ~ dnorm( 0 , 1),
        a[genotype] ~ dnorm(0 , 1)
    ), data=dat_list.Q1 , chains=2 , cores=2 , log_lik=TRUE )
# 
mQ1.a2.error <- ulam(
    alist(
        stem_length ~ dnorm( 1 , p),
        logit(p) <- a*genotype + b*treatment,
        b ~ dnorm( 0 , 1),
        a ~ dnorm(0 , 1)
    ), data=dat_list.Q1error , chains=2 , cores=2 , log_lik=TRUE )
# Do I need to have intercept? 


coeftab(mQ1.a)
coeftab(mQ1.a2.error)
compare(mQ1.a, mQ1.a2.error) # why different?

```

b) include flat without pooling
* I feel I do understand what pooling is meant....
```{r eval=TRUE}
set.seed(15)
mQ1.b <- ulam(
    alist(
        stem_length ~ dnorm( 1 , p),
        logit(p) <- a[genotype] + g[flat_id] + b[treatment],
        b[treatment] ~ dnorm( 0 , sigma_b ),
        a[genotype] ~ dnorm( a_bar , sigma_a ),
        g[flat_id] ~ dnorm( 0 , 1 ),
        a_bar ~ dnorm( 0 , 1.5 ),
        sigma_a ~ dexp(1),
        sigma_b ~ dexp(1)
    ), data=dat_list.Q1 , chains=2 , cores=2 , log_lik=TRUE )
coeftab(mQ1.b)
```

c) use a hierarchical model that allows partial pooling across flats
```{r eval=TRUE}
set.seed(15)
mQ1.c <- ulam(
    alist(
        stem_length ~ dnorm( 1 , p),
        logit(p) <- a[genotype] + g[flat_id] + b[treatment],
        b[treatment] ~ dnorm( 0 , sigma_b ),
        a[genotype] ~ dnorm( a_bar , sigma_a ),
        g[flat_id] ~ dnorm( 0 , sigma_g ),
        a_bar ~ dnorm( 0 , 1.5 ),
        sigma_a ~ dexp(1),
        sigma_g ~ dexp(1),
        sigma_b ~ dexp(1)
    ), data=dat_list.Q1 , chains=2 , cores=2 , log_lik=TRUE )
coeftab(mQ1.c)
```


# Q3) Compare the models, which is preferred?
```{r}
compare(mQ1.a,mQ1.b,mQ1.c)

```

# Q4) Using the hierarchical model, make posterior predictions
a) for average cluster
b) for same clusters
c) showing the "marginal" from cluster
d) showing new clusters.

Q5) Reparameterize the model to help with divergent transitions (even if there aren't any)

Q6--optional)
a) Which genotypes differ from MoneyMaker in Sun conditions?
b) Which genotypes differ from MoneyMaker in Shade conditions?
c) Which genotypes differ from MoneyMaker in their response to shade (difference in sun vs shade)?

# sessionInfo()
```{r}
sessionInfo()
```

