---
title: "Chap14_4"
author: "Kazu"
date: "4/1/2020"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=TRUE,cache = FALSE)
library(rethinking)
library(lme4);library(lmerTest);library(ggplot2);library(reshape2);library(tidyverse);library(readr)
```

# 14.4. Social relations as correlated varying effects
```{r}
data(KosterLeckie) # Gift exchange data
?KosterLeckie
kl_dyads
kl_households
```

```{r R code 14.30}
kl_data <- list(
    N = nrow(kl_dyads),
    N_households = max(kl_dyads$hidB),
    did = kl_dyads$did,
    hidA = kl_dyads$hidA,
    hidB = kl_dyads$hidB,
    giftsAB = kl_dyads$giftsAB,
    giftsBA = kl_dyads$giftsBA
)
m14.4 <- ulam(
    alist(
        giftsAB ~ poisson( lambdaAB ),
        giftsBA ~ poisson( lambdaBA ),
        log(lambdaAB) <- a + gr[hidA,1] + gr[hidB,2] + d[did,1] ,
        log(lambdaBA) <- a + gr[hidB,1] + gr[hidA,2] + d[did,2] ,
        a ~ normal(0,1),
       ## gr matrix of varying effects
       vector[2]:gr[N_households] ~ multi_normal(0,Rho_gr,sigma_gr),
    Rho_gr ~ lkj_corr(4),
    sigma_gr ~ exponential(1),
   ## dyad effects
    transpars> matrix[N,2]:d <-
            compose_noncentered( rep_vector(sigma_d,2) , L_Rho_d , z ),
    matrix[2,N]:z ~ normal( 0 , 1 ),
    cholesky_factor_corr[2]:L_Rho_d ~ lkj_corr_cholesky( 8 ),
    sigma_d ~ exponential(1),
   ## compute correlation matrix for dyads
    gq> matrix[2,2]:Rho_d <<- Chol_to_Corr( L_Rho_d )
), data=kl_data , chains=4 , cores=2 , iter=1000 )
# Warning messages:
# 1: The largest R-hat is NA, indicating chains have not mixed.
# Running the chains for more iterations may help. See
# http://mc-stan.org/misc/warnings.html#r-hat 
# 2: Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.
# Running the chains for more iterations may help. See
# http://mc-stan.org/misc/warnings.html#bulk-ess 
# 3: Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable.
# Running the chains for more iterations may help. See
# http://mc-stan.org/misc/warnings.html#tail-ess 
```

#
```{r R code 14.34}
precis( m14.4 , depth=3 , pars=c("Rho_gr","sigma_gr") )
```

# 
```{r R code 14.35}
dy1 <- apply( post$d[,,1] , 2 , mean )
dy2 <- apply( post$d[,,2] , 2 , mean )
plot( dy1 , dy2 )
```



# 14.5. Continuous categories and the Gaussianprocess
# 14.5.1. Example: Spatial autocorrelation in Oceanic tools.
```{r R code 14.36}
# load the distance matrix
library(rethinking)
data(islandsDistMatrix)
# display (measured in thousands of km)
Dmat <- islandsDistMatrix
colnames(Dmat) <- c("Ml","Ti","SC","Ya","Fi","Tr","Ch","Mn","To","Ha")
round(Dmat,1)
```


```{r R 14.37}
# linear
curve( exp(-1*x) , from=0 , to=4 , lty=2 ,
    xlab="distance" , ylab="correlation" )
# squared
curve( exp(-1*x^2) , add=TRUE )
```

```{r R 14.38}
data(Kline2) # load the ordinary data, now with coordinates
d <- Kline2
d$society <- 1:10 # index observations
dat_list <- list(
    T = d$total_tools,
    P = d$population,
    society = d$society,
    Dmat=islandsDistMatrix )
m14.7 <- ulam(
    alist(
        T ~ dpois(lambda),
        lambda <- (a*P^b/g)*exp(k[society]),
        vector[10]:k ~ multi_normal( 0 , SIGMA ),
        matrix[10,10]:SIGMA <- cov_GPL2( Dmat , etasq , rhosq , 0.01 ),
        c(a,b,g) ~ dexp( 1 ),
        etasq ~ dexp( 2 ),
        rhosq ~ dexp( 0.5 )
    ), data=dat_list , chains=4 , cores=4 , iter=2000 )
```


```{r R 14.39}
precis( m14.7 , depth=3 )
```

```{r R 14.40}
post <- extract.samples(m14.7)
# plot the posterior median covariance function
plot( NULL , xlab="distance (thousand km)" , ylab="covariance" ,
    xlim=c(0,10) , ylim=c(0,2) )
# compute posterior mean covariance
x_seq <- seq( from=0 , to=10 , length.out=100 )
pmcov <- sapply( x_seq , function(x) post$etasq*exp(-post$rhosq*x^2) )
pmcov_mu <- apply( pmcov , 2 , mean )
lines( x_seq , pmcov_mu , lwd=2 )
# plot 60 functions sampled from posterior
for ( i in 1:50 )
    curve( post$etasq[i]*exp(-post$rhosq[i]*x^2) , add=TRUE ,
        col=col.alpha("black",0.3) )
```

# compare models (by Julin, compre chapter 11 model, m11.11)
```{r}

```




```{r R 14.41}
# compute posterior median covariance among societies
K <- matrix(0,nrow=10,ncol=10)
for ( i in 1:10 )
    for ( j in 1:10 )
        K[i,j] <- median(post$etasq) *
                  exp( -median(post$rhosq) * islandsDistMatrix[i,j]^2 )
diag(K) <- median(post$etasq) + 0.01
```

```{r R code 14.42}
# convert to correlation matrix
Rho <- round( cov2cor(K) , 2 )
# add row/col names for convenience
colnames(Rho) <- c("Ml","Ti","SC","Ya","Fi","Tr","Ch","Mn","To","Ha")
rownames(Rho) <- colnames(Rho)
Rho
```


# Error in Rho[i, j] : object of type 'closure' is not subsettable
```{r R code 14.43}
# scale point size to logpop
psize <- d$logpop / max(d$logpop)
psize <- exp(psize*1.5)-2
# plot raw data and labels
plot( d$lon2 , d$lat , xlab="longitude" , ylab="latitude" ,
    col=rangi2 , cex=psize , pch=16 , xlim=c(-50,30) )
labels <- as.character(d$culture)
text( d$lon2 , d$lat , labels=labels , cex=0.7 , pos=c(2,4,3,3,4,1,3,2,4,2) )
# overlay lines shaded by Rho
for( i in 1:10 )
    for ( j in 1:10 )
        if ( i < j )
            lines( c( d$lon2[i],d$lon2[j] ) , c( d$lat[i],d$lat[j] ) ,
                lwd=2 , col=col.alpha("black",Rho[i,j]^2) )
```

```{r R code 14.44}
# compute posterior median relationship, ignoring distance
logpop.seq <- seq( from=6 , to=14 , length.out=30 )
#lambda <- sapply( logpop.seq , function(lp) exp( post$a + post$bp*lp ) )
# by Julin
lambda <- sapply( logpop.seq , function(lp) post$a*exp(lp)^post$b/post$g )

lambda.median <- apply( lambda , 2 , median )
lambda.PI80 <- apply( lambda , 2 , PI , prob=0.8 )
# plot raw data and labels
plot( d$logpop , d$total_tools , col=rangi2 , cex=psize , pch=16 ,
    xlab="log population" , ylab="total tools" )
text( d$logpop , d$total_tools , labels=labels , cex=0.7 ,
    pos=c(4,3,4,2,2,1,4,4,4,2) )
# display posterior predictions
lines( logpop.seq , lambda.median , lty=2 )
lines( logpop.seq , lambda.PI80[1,] , lty=2 )
lines( logpop.seq , lambda.PI80[2,] , lty=2 )
# overlay correlations
for( i in 1:10 )
    for ( j in 1:10 )
        if ( i < j )
            lines( c( d$logpop[i],d$logpop[j] ) ,
                   c( d$total_tools[i],d$total_tools[j] ) ,
                   lwd=2 , col=col.alpha("black",Rho[i,j]^2) )
```

```{r}

```

```{r R code 14.45}
m14.7nc <- ulam(
    alist(
        T ~ dpois(lambda),
        lambda <- (a*P^b/g)*exp(k[society]),
        # non-centered Gaussian Process prior
        transpars> vector[10]: k <<- L_SIGMA * z,
        vector[10]: z ~ normal( 0 , 1 ),
        transpars> matrix[10,10]: L_SIGMA <<- cholesky_decompose( SIGMA ),
        transpars> matrix[10,10]: SIGMA <- cov_GPL2( Dmat , etasq , rhosq , 0.01 ),
        c(a,b,g) ~ dexp( 1 ),
        etasq ~ dexp( 2 ),
        rhosq ~ dexp( 0.5 )
    ), data=dat_list , chains=4 , cores=4 , iter=2000 )
```

# 14.5.2. Example: Phylogenetic distance.
```{r}
library(rethinking)
data(Primates301)
data(Primates301_nex)
# plot it
library(ape)
### cut in a book. Use code in github
plot( ladderize(Primates301_nex) , type="fan" , font=1 , no.margin=TRUE ,
    label.offset=1 , cex=0.5 )

```

```{r}
d <- Primates301
d$name <- as.character(d$name)
dstan <- d[ complete.cases( d$group_size , d$body , d$brain ) , ]
spp_obs <- dstan$name
```

```{r}
dat_list <- list(
    N_spp = nrow(dstan),
    M = standardize(log(dstan$body)),
    B = standardize(log(dstan$brain)),
    G = standardize(log(dstan$group_size)),
    Imat = diag( nrow(dstan) )
)
```



```{r}
m14.8 <- ulam(
    alist(
        G ~ multi_normal( mu , SIGMA ),
        mu <- a + bM*M + bB*B,
        matrix[N_spp,N_spp]: SIGMA <- Imat * sigma_sq,
        a ~ normal( 0 , 1 ),
        c(bM,bB) ~ normal( 0 , 0.5 ),
        sigma_sq ~ exponential( 1 )
    ), data=dat_list , chains=4 , cores=4 )
precis( m14.8 )
```

```{r R code 14.49}
tree_trimmed <- keep.tip( Primates301_nex, spp_obs )
Rbm <- corBrownian( phy=tree_trimmed )
V <- vcv(Rbm)
Dmat <- cophenetic( tree_trimmed )
plot( Dmat , V , xlab="phylogenetic distance" , ylab="covariance" )
```

```{r R code 14.50}
# put species in right order
dat_list$V <- V[ spp_obs , spp_obs ]
# convert to correlation matrix
dat_list$R <- dat_list$V / max(V)
# Brownian motion model
m14.9 <- ulam(
    alist(
        G ~ multi_normal( mu , SIGMA ),
        mu <- a + bM*M + bB*B,
        matrix[N_spp,N_spp]: SIGMA <- R * sigma_sq,
        a ~ normal( 0 , 1 ),
        c(bM,bB) ~ normal( 0 , 0.5 ),
        sigma_sq ~ exponential( 1 )
    ), data=dat_list , chains=4 , cores=4 )
precis( m14.9 )
# heatmap (from book text)
image(V)
image(Dmat)
```

```{r R code 14.51}
# add scaled and reordered distance matrix
dat_list$Dmat <- Dmat[ spp_obs , spp_obs ] / max(Dmat)
m14.10 <- ulam(
    alist(
        G ~ multi_normal( mu , SIGMA ),
        mu <- a + bM*M + bB*B,
        matrix[N_spp,N_spp]: SIGMA <- cov_GPL2( Dmat , etasq , rhosq , 0.01 ),
        a ~ normal(0,1),
        c(bM,bB) ~ normal(0,0.5),
        etasq ~ exponential(1),
        rhosq ~ exponential(1)
    ), data=dat_list , chains=4 , cores=4 )
precis( m14.10 )
```

```{r R code 14.52}
 post <- extract.samples(m14.10)
plot( NULL , xlim=c(0,max(dat_list$Dmat)) , ylim=c(0,5) ,
    xlab="phylogenetic distance" , ylab="covariance" )
for ( i in 1:50 ) curve( post$etasq[i]*exp(-post$rhosq[i]*x^2) , add=TRUE , col=grau() )
```


# 14.7. Practice
# 14M5. 
* Modify the phylogenetic distance example to use group size as the outcome and brain size as
a predictor. Assuming brain size influences group size, what is your estimate of the effect? How does
phylogeny influence the estimate? 
```{r}

```


# 14H2
* which is 234 height measurements on 26 boys from an Oxford Boys Club (I think these were like youth athletic leagues?), at 9 different ages (centered and standardized) per boy. You’ll be interested in predicting height, using age, clustered by Subject (individual boy).
* Fit a model with varying intercepts and slopes (on age), clustered by Subject. Present and interpret the parameter estimates. Which varying effect contributes more variation to the heights, the intercept or the slope?
```{r}
data(Oxboys)
summary(Oxboys)
head(Oxboys)
Oxboys$Subject <- as.character(Oxboys$Subject)

# looking raw data
Oxboys %>% ggplot(aes(x=age, y=height,color=Subject)) + geom_line()
```

# 
```{r}
Oxboys.list <- list(
  Subject=as.integer(Oxboys$Subject),
  age=Oxboys$age,
  height=Oxboys$height
)
mOxboys.partialpooling <- ulam(
    alist(
        height ~ dnorm( mu , sigma ),
        mu <- a[Subject] + b[Subject]*age, 
        a[Subject] ~ dnorm(a_bar,a_sigma),
        b[Subject] ~ dnorm(0,1),  
        a_bar ~ dnorm(145,1),
        a_sigma ~ dexp(1),
        sigma ~ dexp(1)
    ) , data=Oxboys.list , chains=4 , cores=2,iter=1000,log_lik=TRUE)
# 
precis( mOxboys.partialpooling )
plot(mOxboys.partialpooling)
```

```{r}
mOxboys.multinormal <- ulam(
  alist(
    height ~ normal(mu, sigma),
    mu <- a[Subject] + bS[Subject] * age,
    c(a, bS)[Subject] ~ multi_normal(c(a_S, b_S) , Rho , sigma_S),
    a_S ~ normal(145, 1),
    b_S ~ normal(0, 1),
    sigma_S ~ exponential(1),
    sigma ~ exponential(1),
    Rho ~ lkj_corr(2)
  ),
  data = Oxboys.list,
  chains = 4,
  cores = 4,
  log_lik = T,
  iter = 5000
)
precis(mOxboys.multinormal, depth = 3, pars = c("a", "bS", "Rho", "sigma"))
```


# 14H3
* Now consider the correlation between the varying intercepts and slopes. 
* Can you explain its value? How would this estimated correlation influence your predictions about a new sample of boys?
```{r}


```


# 14H3
```{r}

```




```{r}
sessionInfo()
```

