---
title: "Chap14"
author: "Kazu"
date: "3/5/2020"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=TRUE,cache = FALSE)
library(rethinking)
library(lme4);library(lmerTest);library(ggplot2);library(reshape2);library(tidyverse);library(readr)
```

# 14.1. Varying slopes by construction
## 14.1.1. Simulate the population. 
```{r R code 14.1}
a <- 3.5
b <- (-1)
sigma_a <- 1
sigma_b <- 0.5
rho <- (-0.7)
# average morning wait time
# average difference afternoon wait time
# std dev in intercepts
# std dev in slopes
# correlation between intercepts and slopes
```

```{r}
Mu <- c( a , b )
```

```{r}
cov_ab <- sigma_a*sigma_b*rho
Sigma <- matrix( c(sigma_a^2,cov_ab,cov_ab,sigma_b^2) , ncol=2 )
```

```{r}
 matrix( c(1,2,3,4) , nrow=2 , ncol=2 )
```

```{r}
sigmas <- c(sigma_a,sigma_b) # standard deviations
Rho <- matrix( c(1,rho,rho,1) , nrow=2 ) # correlation matrix
# now matrix multiply to get covariance matrix
Sigma <- diag(sigmas) %*% Rho %*% diag(sigmas)
```

```{r}
 N_cafes <- 20
```

```{r}
library(MASS)
set.seed(5) # used to replicate example
vary_effects <- mvrnorm( N_cafes , Mu , Sigma )
```

```{r}
a_cafe <- vary_effects[,1]
b_cafe <- vary_effects[,2]
```

```{r}
plot( a_cafe , b_cafe , col=rangi2 ,
    xlab="intercepts (a_cafe)" , ylab="slopes (b_cafe)" )
# overlay population distribution
library(ellipse)
for ( l in c(0.1,0.3,0.5,0.8,0.99) )
    lines(ellipse(Sigma,centre=Mu,level=l),col=col.alpha("black",0.2))
```

# 14.1.2. Simulate observations.
```{r}
set.seed(22)
N_visits <- 10
afternoon <- rep(0:1,N_visits*N_cafes/2)
cafe_id <- rep( 1:N_cafes , each=N_visits )
mu <- a_cafe[cafe_id] + b_cafe[cafe_id]*afternoon
sigma <- 0.5  # std dev within cafes
wait <- rnorm( N_visits*N_cafes , mu , sigma )
d <- data.frame( cafe=cafe_id , afternoon=afternoon , wait=wait )
```

# 14.1.3. The varying slopes model.
```{r}
R <- rlkjcorr( 1e4 , K=2 , eta=2 )
dens( R[,1,2] , xlab="correlation" )
```

```{r}
m14.1 <- ulam(
    alist(
        wait ~ normal( mu , sigma ),
        mu <- a_cafe[cafe] + b_cafe[cafe]*afternoon,
        c(a_cafe,b_cafe)[cafe] ~ multi_normal( c(a,b) , Rho , sigma_cafe ),
        a ~ normal(5,2),
        b ~ normal(-1,0.5),
        sigma_cafe ~ exponential(1),
        sigma ~ exponential(1),
        Rho ~ lkj_corr(2)
    ) , data=d , chains=4 , cores=2,iter=2000,log_lik=TRUE)
# What is lkj_corr?
?rlkjcorr
?rmultinom # for multi_ormal, correct?
```

```{r}
post <- extract.samples(m14.1)
dens( post$Rho[,1,2] )
```

```{r R code 14.14}
# compute unpooled estimates directly from data
a1 <- sapply( 1:N_cafes ,
        function(i) mean(wait[cafe_id==i & afternoon==0]) )
b1 <- sapply( 1:N_cafes ,
        function(i) mean(wait[cafe_id==i & afternoon==1]) ) - a1
# extract posterior means of partially pooled estimates
post <- extract.samples(m14.1) # m14.1 instead of m13.1?
a2 <- apply( post$a_cafe , 2 , mean )
b2 <- apply( post$b_cafe , 2 , mean )
# plot both and connect with lines
plot( a1 , b1 , xlab="intercept" , ylab="slope" ,
    pch=16 , col=rangi2 , ylim=c( min(b1)-0.1 , max(b1)+0.1 ) ,
    xlim=c( min(a1)-0.1 , max(a1)+0.1 ) )
points(a2,b2,pch=1) 
for ( i in 1:N_cafes ) lines( c(a1[i],a2[i]) , c(b1[i],b2[i]))

```

```{r R 14.15}
# compute posterior mean bivariate Gaussian
Mu_est <- c( mean(post$a) , mean(post$b) )
rho_est <- mean( post$Rho[,1,2] )
sa_est <- mean( post$sigma_cafe[,1] )
sb_est <- mean( post$sigma_cafe[,2] )
cov_ab <- sa_est*sb_est*rho_est
Sigma_est <- matrix( c(sa_est^2,cov_ab,cov_ab,sb_est^2) , ncol=2 )
# draw contours
library(ellipse)
for ( l in c(0.1,0.3,0.5,0.8,0.99) )
    lines(ellipse(Sigma_est,centre=Mu_est,level=l),
        col=col.alpha("black",0.2))
```

* outcome scale
```{r R code 14.16}
# convert varying effects to waiting times
wait_morning_1 <- (a1)
wait_afternoon_1 <- (a1 + b1)
wait_morning_2 <- (a2)
wait_afternoon_2 <- (a2 + b2)
# plot both and connect with lines
plot( wait_morning_1 , wait_afternoon_1 , xlab="morning wait" ,
    ylab="afternoon wait" , pch=16 , col=rangi2 ,
    ylim=c( min(wait_afternoon_1)-0.1 , max(wait_afternoon_1)+0.1 ) ,
    xlim=c( min(wait_morning_1)-0.1 , max(wait_morning_1)+0.1 ) )
points( wait_morning_2 , wait_afternoon_2 , pch=1 )
for ( i in 1:N_cafes )
    lines( c(wait_morning_1[i],wait_morning_2[i]) ,
    c(wait_afternoon_1[i],wait_afternoon_2[i]) )
abline( a=0 , b=1 , lty=2 )
```

```{r R cpde 14.17}
# now shrinkage distribution by simulation
v <- mvrnorm( 1e4 , Mu_est , Sigma_est ) # mvrnorm()
v[,2] <- v[,1] + v[,2] # calculate afternoon wait
Sigma_est2 <- cov(v)
Mu_est2 <- Mu_est
Mu_est2[2] <- Mu_est[1]+Mu_est[2]
# draw contours
library(ellipse)
for ( l in c(0.1,0.3,0.5,0.8,0.99) )
    lines(ellipse(Sigma_est2,centre=Mu_est2,level=l),
        col=col.alpha("black",0.5))
```



# problems
14E1

* original

$$
y_i \sim Normal(\mu_i, \sigma) \\
\mu_i = \alpha_{GROUP[i]} + \beta x_i \\
\alpha_{GROUP} \sim Normal(\alpha, \sigma_\alpha) \\
\alpha \sim Normal(0, 10) \\
\beta \sim Normal(0, 1) \\ 
\sigma \sim HalfCauchy(0,2) \\ 
\sigma_\alpha \sim HalfCouchy(0, 2)
$$

* VARYING SLOPES ON THE PREDICTOR X

yi ~ Normal(mu_i, sigma)
mu_i = alpha_group[i] + beta[i] * xi
[alpha_group[i], beta[i]] ~ MVNormal([alpha, beta], S)
S = (sigma_alpha, 0, 0, sigma_beta)_2x2 R (sigma_alpha, 0, 0, sigma_beta)_2x2
alpha ~ Normal(0, 10)
beta ~ Normal(0, 1)
sigma ~ HalfCauchy(0, 2)
sigma_alpha ~ HalfCauchy(0, 2)
sigma_beta ~ HalfCauchy(0, 2)
R ~ LKJcorr(2)


* I gave up... 
$$
y_i \sim Normal(\mu_i, \ sigma) \\
\mu_i = \alpha_{GROUP[i]} + \beta x_i \\
\alpha_{GROUP} \sim Normal(\alpha, \sigma_\alpha) \\

\begin{bmatrix}\alpha_{GROUP} \\
\beta
\end{bmatrix} \sim MVNormal\
begin{pmatrix}
  \begin{bmatrix}\alpha_{GROUP} \\
  \beta
  \end{bmatrix}
  & S
\end{pmatrix}\\

S = \begin{pmatrix}a & b\\\ c & d\end{pmatrix} R \begin{pmatrix}a & b\\\ c & d\end{pmatrix} \\\\

\alpha \sim Normal(0, 10) \\
\beta \sim Normal(0, 1) \\ 
\sigma \sim HalfCauchy(0,2) \\ 
\sigma_\alpha \sim HalfCouchy(0, 2) \\


X = \begin{bmatrix}1 & x_{1}\\
1 & x_{2}\\
1 & x_{3}
\end{bmatrix} \\

S = \begin{pmatrix}a & b\\\ c & d\end{pmatrix} \\
$$


14E2 Think up a context in which varying intercepts will be positively correlated with varying
slopes. Provide a mechanistic explanation for the correlation.
(Kazu's answer) Taller swimmers are heavyer. Tall and heavy swimmers swim faster because they can pull more water per stroke than shorter swimmers do and their stroke speed is maintained by more muscles.

14E3 When is it possible for a varying slopes model to have fewer effective parameters (asestimated by WAIC or DIC) than the corresponding model with fixed (unpooled) slopes? Explain.
```{r}

```

14M1. Repeat the cafe robot simulation from the beginning of the chapter. This time, set rho to zero, so that there is no correlation between intercepts and slopes. How does the posterior distribution of the correlation reflect this change in the underlying simulation
```{r}
a <- 3.5
b <- (-1)
sigma_a <- 1
sigma_b <- 0.5
# rho <- (-0.7)
rho <- 0 # new 
# average morning wait time
# average difference afternoon wait time
# std dev in intercepts
# std dev in slopes
# correlation between intercepts and slopes
cov_ab <- sigma_a*sigma_b*rho
Sigma <- matrix( c(sigma_a^2,cov_ab,cov_ab,sigma_b^2) , ncol=2 )
sigmas <- c(sigma_a,sigma_b) # standard deviations
Rho <- matrix( c(1,rho,rho,1) , nrow=2 ) # correlation matrix
# now matrix multiply to get covariance matrix
Sigma <- diag(sigmas) %*% Rho %*% diag(sigmas)
 N_cafes <- 20
library(MASS)
set.seed(5) # used to replicate example
vary_effects <- mvrnorm( N_cafes , Mu , Sigma )
a_cafe <- vary_effects[,1]
b_cafe <- vary_effects[,2]
#
plot( a_cafe , b_cafe , col=rangi2 ,
    xlab="intercepts (a_cafe)" , ylab="slopes (b_cafe)" )
# overlay population distribution
library(ellipse)
for ( l in c(0.1,0.3,0.5,0.8,0.99) )
    lines(ellipse(Sigma,centre=Mu,level=l),col=col.alpha("black",0.2))
#
set.seed(22)
N_visits <- 10
afternoon <- rep(0:1,N_visits*N_cafes/2)
cafe_id <- rep( 1:N_cafes , each=N_visits )
mu <- a_cafe[cafe_id] + b_cafe[cafe_id]*afternoon
sigma <- 0.5  # std dev within cafes
wait <- rnorm( N_visits*N_cafes , mu , sigma )
d <- data.frame( cafe=cafe_id , afternoon=afternoon , wait=wait )

#
R <- rlkjcorr( 1e4 , K=2 , eta=2 )
dens( R[,1,2] , xlab="correlation" )
```
* model
```{r}
m14.1.rhozero <- ulam(
    alist(
        wait ~ normal( mu , sigma ),
        mu <- a_cafe[cafe] + b_cafe[cafe]*afternoon,
        c(a_cafe,b_cafe)[cafe] ~ multi_normal( c(a,b) , Rho , sigma_cafe ),
        a ~ normal(5,2),
        b ~ normal(-1,0.5),
        sigma_cafe ~ exponential(1),
        sigma ~ exponential(1),
        Rho ~ lkj_corr(2)
    ) , data=d , chains=4 , cores=2,iter=2000,log_lik=TRUE)

```
* posterior distribution

```{r}
post <- extract.samples(m14.1.rhozero)
dens( post$Rho[,1,2] )
#
# compute unpooled estimates directly from data
a1 <- sapply( 1:N_cafes ,
        function(i) mean(wait[cafe_id==i & afternoon==0]) )
b1 <- sapply( 1:N_cafes ,
        function(i) mean(wait[cafe_id==i & afternoon==1]) ) - a1
# extract posterior means of partially pooled estimates
post <- extract.samples(m14.1.rhozero) # m14.1 instead of m13.1?
a2 <- apply( post$a_cafe , 2 , mean )
b2 <- apply( post$b_cafe , 2 , mean )
# plot both and connect with lines
plot( a1 , b1 , xlab="intercept" , ylab="slope" ,
    pch=16 , col=rangi2 , ylim=c( min(b1)-0.1 , max(b1)+0.1 ) ,
    xlim=c( min(a1)-0.1 , max(a1)+0.1 ) )
points(a2,b2,pch=1) 
for ( i in 1:N_cafes ) lines( c(a1[i],a2[i]) , c(b1[i],b2[i]))
```

14M2
Fit this multilevel model to the simulated café data:
Wi ∼ Normal(μi, σ)
μi = αcafé[i] + βcafé[i]A
i αcafé ∼ Normal(α, σα)
βcafé ∼Normal(β,σβ)
α ∼ Normal(0, 10) # should be Normal(5,2)?
β ∼ Normal(0, 10) # should be Normal(-1,0.5)?
σ ∼ HalfCauchy(0, 1)
σα ∼ HalfCauchy(0, 1) σβ ∼HalfCauchy(0,1)

```{r}
# I do not understand how to use HalfCauchy distribution function
library(extraDistr)
dhcauchy(0,1) # errr

m14.1.problemM2 <- ulam(
    alist(
        wait ~ normal( mu , sigma ),
        mu <- a[cafe] + b[cafe]*afternoon,
        a[cafe] ~ normal(alpha,sigma_alpha),
        b[cafe] ~ normal(beta,sigma_beta),
        alpha ~ normal(5,2),
        beta ~ normal(-1,0.5),
        sigma ~  exponential(1),
        sigma_alpha ~ exponential(1),
        sigma_beta ~ exponential(1)
    ) , data=d , chains=4 , cores=2,iter=2000,log_lik=TRUE)

precis(m14.1.problemM2, depth=2 )
plot( precis(m14.1.problemM2,depth=2) )
```
# compare
```{r}
compare(m14.1,m14.1.problemM2)
compare(m14.1,m14.1.rhozero,m14.1.problemM2)

```

