---
title: "Chap9"
author: "Kazu"
date: "9/27/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=TRUE)
library(rethinking)
library(lme4);library(lmerTest);library(ggplot2);library(reshape2);library(tidyverse);library(readr)
```

# Quad approximation does not work with non Gausian distribution... (by J)

```{r}
num_weeks <- 1e5 
positions <- rep(0,num_weeks) 
current <- 10
for ( i in 1:num_weeks ) { 
  # record current position 
  positions[i] <- current 
  # flip coin to generate proposal 
  proposal <- current + sample( c(-1,1) , size=1 ) 
  # now make sure he loops around the archipelago 
  if ( proposal < 1 ) proposal <- 10 
  if ( proposal > 10 ) proposal <- 1 
  # move? 
  prob_move <- proposal/current 
  current <- ifelse( runif(1) < prob_move , proposal , current ) 
}
plot(x=1:1000, y=positions[1:1000])
```

## R code 9.2
```{r}
D <- 10
T <- 1e3
Y <- rmvnorm(T,rep(0,D),diag(D))
rad_dist <- function( Y ) sqrt( sum(Y^2) )
Rd <- sapply( 1:T , function(i) rad_dist( Y[i,] ) )
dens( Rd )
```
## R code 9.3
```{r}
# U needs to return neg-log-probability
myU4 <- function( q , a=0 , b=1 , k=0 , d=1 ) {
  muy <- q[1]
  mux <- q[2]
  U <- sum( dnorm(y,muy,1,log=TRUE) ) + sum( dnorm(x,mux,1,log=TRUE) ) +
    dnorm(muy,a,b,log=TRUE) + dnorm(mux,k,d,log=TRUE)
  return( -U )
}
```
## R code 9.4
```{r}
# gradient function
# need vector of partial derivatives of U with respect to vector q
myU_grad4 <- function( q , a=0 , b=1 , k=0 , d=1 ) {
  muy <- q[1]
  mux <- q[2]
  G1 <- sum( y - muy ) + (a - muy)/b^2 #dU/dmuy
  G2 <- sum( x - mux ) + (k - mux)/d^2 #dU/dmuy
  return( c( -G1 , -G2 ) ) # negative bc energy is neg-log-prob
}
# test data
set.seed(7)
y <- rnorm(50)
x <- rnorm(50)
x <- as.numeric(scale(x))
y <- as.numeric(scale(y))
```
## R code 9.5
```{r}
library(shape) # for fancy arrows
Q <- list()
Q$q <- c(-0.1,0.2)
pr <- 0.3
plot( NULL , ylab="muy" , xlab="mux" , xlim=c(-pr,pr) , ylim=c(-pr,pr) )
step <- 0.03
L <- 11 # 0.03/28 for U-turns --- 11 for working example
n_samples <- 4
path_col <- col.alpha("black",0.5)
points( Q$q[1] , Q$q[2] , pch=4 , col="black" )
for ( i in 1:n_samples ) {
  Q <- HMC2( myU4 , myU_grad4 , step , L , Q$q )
  if ( n_samples < 10 ) {
    for ( j in 1:L ) {
      K0 <- sum(Q$ptraj[j,]^2)/2 # kinetic energy
      lines( Q$traj[j:(j+1),1] , Q$traj[j:(j+1),2] , col=path_col , lwd=1+2*K0 )
    }
    points( Q$traj[1:L+1,] , pch=16 , col="white" , cex=0.35 )
    Arrows( Q$traj[L,1] , Q$traj[L,2] , Q$traj[L+1,1] , Q$traj[L+1,2] ,
            arr.length=0.35 , arr.adj = 0.7 )
    text( Q$traj[L+1,1] , Q$traj[L+1,2] , i , cex=0.8 , pos=4 , offset=0.4 )
  }
  points( Q$traj[L+1,1] , Q$traj[L+1,2] , pch=ifelse( Q$accept==1 , 16 , 1 ) ,
          col=ifelse( abs(Q$dH)>0.1 , "red" , "black" ) )
}
```
## R code 9.6
```{r}
HMC2 <- function (U, grad_U, epsilon, L, current_q) {
  q = current_q
  p = rnorm(length(q),0,1) # random flick - p is momentum.
  current_p = p
  # Make a half step for momentum at the beginning
  p = p - epsilon * grad_U(q) / 2
  # initialize bookkeeping - saves trajectory
  qtraj <- matrix(NA,nrow=L+1,ncol=length(q))
  ptraj <- qtraj
  qtraj[1,] <- current_q
  ptraj[1,] <- p
  
  ## R code 9.7
  # Alternate full steps for position and momentum
  for ( i in 1:L ) {
    q = q + epsilon * p # Full step for the position
    # Make a full step for the momentum, except at end of trajectory
    if ( i!=L ) {
      p = p - epsilon * grad_U(q)
      ptraj[i+1,] <- p
    }
    qtraj[i+1,] <- q
  }
  ## R code 9.8
  # Make a half step for momentum at the end
  p = p - epsilon * grad_U(q) / 2
  ptraj[L+1,] <- p
  # Negate momentum at end of trajectory to make the proposal symmetric
  p = -p
  # Evaluate potential and kinetic energies at start and end of trajectory
  current_U = U(current_q)
  current_K = sum(current_p^2) / 2
  proposed_U = U(q)
  proposed_K = sum(p^2) / 2
  # Accept or reject the state at end of trajectory, returning either
  # the position at the end of the trajectory or the initial position
  accept <- 0
  if (runif(1) < exp(current_U-proposed_U+current_K-proposed_K)) {
    new_q <- q  # accept
    accept <- 1
  } else new_q <- current_q  # reject
  return(list( q=new_q, traj=qtraj, ptraj=ptraj, accept=accept ))
}
```
# 9.4. EasyHMC:ulam
## the terrain ruggedness example from Chapter 7. This code will load the data and reduce it down to cases (nations) that have the outcome variable of interest:

## R code 9.9
```{r}
library(rethinking)
data(rugged)
d <- rugged
d$log_gdp <- log(d$rgdppc_2000) # log GDP
# extract countries with GDP data
dd <- d[ complete.cases(d$rgdppc_2000) , ] 
# rescale variables
dd$log_gdp_std <- dd$log_gdp / mean(dd$log_gdp)
dd$rugged_std <- dd$rugged / max(dd$rugged)
# make variable to index Africa (1) or not (2) (pg 249)
dd$cid <- ifelse( dd$cont_africa==1 , 1 , 2 )
```
## R code 9.10 (the old way)
```{r}  
m8.5 <- quap(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp( 1 )
  ) ,
  data=dd )
precis( m8.5 , depth=2 )
```
## R code 9.11
```{r}
dat_slim <- list(
  log_gdp_std = dd$log_gdp_std,
  rugged_std = dd$rugged_std,
  cid = as.integer( dd$cid )
)
str(dat_slim)
```
# 9.4.2. Sampling from the posterior
## R code 9.12
```{r}
m9.1 <- ulam(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp(1)
  ) ,
  data=dat_slim , chains=1 )
```
## R code 9.13
```{r}
precis( m9.1 , depth=2 )
```
# 9.4.3. Samplingagain,inparallel.
## R code 9.14
```{r}
m9.1 <- ulam(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp( 1 )
  ) ,
  data=dat_slim , chains=4 , cores=2 , iter=1000 )
```
## R code 9.15
```{r}
show( m9.1 )
```
## R code 9.16
```{r}
precis( m9.1 , 2 )
```
# 9.4.4. Visualization
## R code 9.17
```{r}
pairs( m9.1 )
```
# 9.4.5. Checking the chain.
## R code 9.18
```{r}
traceplot( m9.1 )
```
## R code 9.19
```{r}
trankplot( m9.1 , n_cols=2 )
```

# 9.7 Practice #####
## 8E3. Which sort of parameters can Hamiltonian Monte Carlo not handle? Can you explain why?
## (A) discrete

## 8M1. Re-estimate the terrain ruggedness model from the chapter,but now using a uniform prior and an exponential prior for the standard deviation, sigma. The uniform prior should be dunif(0,10) and the exponential should be dexp(1). Do the different priors have any detectible influence on the posterior distribution?
# 9.4.2. Sampling from the posterior
## R code 9.12
```{r}

# uniform prior for standar deviation
m9.1.sigma.unif <- ulam(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dunif(0,10) # uniform prior
  ) ,
  data=dat_slim , chains=2,cores=2 ) # try chains 2 and cores 2
# exponential for standar deviation
m9.1.sigma.expo <- ulam(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp(1) # an exponential prior
  ) ,
  data=dat_slim , chains=2, cores=2 )

```
# posterior
```{r}
precis( m9.1.sigma.unif , depth=2 )
precis( m9.1.sigma.expo , depth=2 )
rethinking::compare(m9.1.sigma.unif,m9.1.sigma.expo)
```

## 8M2. 
* The Cauchy and exponential priors from the terrain ruggedness model are very weak. They can be made more informative by reducing their scale. Compare the dcauchy and dexp priors for progressively smaller values of the scaling parameter. As these priors become stronger, how does each influence the posterior distribution?

```{r}
# Cauchy distribution
m9.1.sigma.cauchy.1 <- ulam(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dcauchy(0,1) # cauchy distribution
  ) ,
  data=dat_slim , chains=2,cores=2,log_lik = TRUE)
m9.1.sigma.cauchy.0.1 <- ulam(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dcauchy(0,0.1) # cauchy distribution
  ) ,
  data=dat_slim , chains=2,cores=2,log_lik = TRUE )
m9.1.sigma.cauchy.0.01 <- ulam(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dcauchy(0,0.01) # cauchy distribution
  ) ,
  data=dat_slim , chains=2,cores=2,log_lik = TRUE )
m9.1.sigma.cauchy.0.001 <- ulam(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dcauchy(0,0.001) # cauchy distribution
  ) ,
  data=dat_slim , chains=2,cores=2,log_lik = TRUE )
# exponential distribution for standar deviation
m9.1.sigma.expo.1 <- ulam(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp(1) # an exponential prior
  ) ,
  data=dat_slim , chains=2, cores=2,log_lik = TRUE )
m9.1.sigma.expo.0.1 <- ulam(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp(0.1) # an exponential prior
  ) ,
  data=dat_slim , chains=2, cores=2,log_lik = TRUE )
m9.1.sigma.expo.0.01 <- ulam(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp(0.01) # an exponential prior
  ) ,
  data=dat_slim , chains=2, cores=2,log_lik = TRUE )
m9.1.sigma.expo.0.001 <- ulam(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp(0.001) # an exponential prior
  ) ,
  data=dat_slim , chains=2, cores=2,log_lik = TRUE )
```
#
```{r}
precis(m9.1.sigma.cauchy.1,depth=2)
precis(m9.1.sigma.cauchy.0.1,depth=2)
precis(m9.1.sigma.cauchy.0.01,depth=2)
precis(m9.1.sigma.cauchy.0.001,depth=2)
precis(m9.1.sigma.expo.1,depth=2)
precis(m9.1.sigma.expo.0.1,depth=2)
precis(m9.1.sigma.expo.0.01,depth=2)
precis(m9.1.sigma.expo.0.001,depth=2)
compare(m9.1.sigma.cauchy.1,m9.1.sigma.cauchy.0.1,m9.1.sigma.cauchy.0.01,m9.1.sigma.cauchy.0.001)
compare(m9.1.sigma.expo.1,m9.1.sigma.expo.0.1,m9.1.sigma.expo.0.01,m9.1.sigma.expo.0.001)
```


## 8H1. 
* Run the model below and then inspect the posterior distribution and explain what it is accomplishing.
* model
```{r error=TRUE}
mp <- map2stan(
    alist(
        a ~ dnorm(0,1),
        b ~ dcauchy(0,1)
    ),
    data=list(y=1),
    start=list(a=0,b=0),
    iter=1e4, warmup=100 , WAIC=FALSE,chains=2,cores=2)
```
# 
```{r}
precis(mp,depth=2)
```
#
```{r}
traceplot(mp)
```
# 
```{r}
trankplot(mp)
```



### Compare the samples for the parameters a and b. Can you explain the different trace plots, using what you know about the Cauchy distribution?


## 8H2. Recall the divorce rate example from Chapter 5. Repeat that analysis, using map2stan this time, fitting models m5.1, m5.2, and m5.3. Use compare to compare the models on the basis of WAIC. Explain the results.
```{r}
# load data and copy
#library(rethinking)
data(WaffleDivorce)
d <- WaffleDivorce
# standardize variables
d$A <- scale( d$MedianAgeMarriage )
d$D <- scale( d$Divorce )
```

```{r}
m5.1 <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bA * A ,
        a ~ dnorm( 0 , 0.2 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
) , data = d )
```

```{r}
d$M <- scale( d$Marriage )
m5.2 <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bM * M ,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
) , data = d )
```

```{r}
m5.3 <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bM*M + bA*A ,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
```

