---
title: "Chap14_2_HW"
author: "Kazu"
date: "3/13/2020"
output:
  github_document:
    pandoc_args: --webtex
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=TRUE,cache = FALSE)
library(rethinking)
library(lme4);library(lmerTest);library(ggplot2);library(reshape2);library(tidyverse);library(readr)
```

# test (my rstan had errors so I checked rstan with simple sample from book)
# multilevel tadpoles
```{r}
data(reedfrogs)
d <- reedfrogs
str(d)
d
# make the tank cluster variable
d$tank <- 1:nrow(d)
dat <- list(
    S = d$surv,
    N = d$density,
    tank = d$tank )

```

* varying intercepts model (= simplelst kind of varying effects) 

$$
Si \sim Bionial(N_i,p_i) \\
logit(p_i) \sim \alpha_{TANK[i]} \\
\alpha_j \sim Normal(0,1.5),\ for\ j=1..48 \\
$$

```{r}
m13.1 <- ulam(
    alist(
        S ~ dbinom( N , p ),
        logit(p) <- a[tank],
        a[tank] ~ dnorm( 0 , 1.5 )
), data=dat, chains=4, log_lik=TRUE ,cores=2) # cores=2 added by Kazu
precis( m13.1 , depth=2 )

```

# the end of test1

# varying slope test (varying slope version of practice 13M3, (eg. m.practice13M3.b) had error, so I cheched varying slope codes from book)
## 14.1.1. Simulate the population. 
```{r R code 14.1}
a <- 3.5
b <- (-1)
sigma_a <- 1
sigma_b <- 0.5
rho <- (-0.7)
# average morning wait time
# average difference afternoon wait time
# std dev in intercepts
# std dev in slopes
# correlation between intercepts and slopes
```

```{r}
Mu <- c( a , b )
```

```{r}
cov_ab <- sigma_a*sigma_b*rho
Sigma <- matrix( c(sigma_a^2,cov_ab,cov_ab,sigma_b^2) , ncol=2 )
```

```{r}
 matrix( c(1,2,3,4) , nrow=2 , ncol=2 )
```

```{r}
sigmas <- c(sigma_a,sigma_b) # standard deviations
Rho <- matrix( c(1,rho,rho,1) , nrow=2 ) # correlation matrix
# now matrix multiply to get covariance matrix
Sigma <- diag(sigmas) %*% Rho %*% diag(sigmas)
```

```{r}
 N_cafes <- 20
```

```{r}
library(MASS)
set.seed(5) # used to replicate example
vary_effects <- mvrnorm( N_cafes , Mu , Sigma )
```

```{r}
a_cafe <- vary_effects[,1]
b_cafe <- vary_effects[,2]
```

```{r}
plot( a_cafe , b_cafe , col=rangi2 ,
    xlab="intercepts (a_cafe)" , ylab="slopes (b_cafe)" )
# overlay population distribution
library(ellipse)
for ( l in c(0.1,0.3,0.5,0.8,0.99) )
    lines(ellipse(Sigma,centre=Mu,level=l),col=col.alpha("black",0.2))
```

# 14.1.2. Simulate observations.
```{r}
set.seed(22)
N_visits <- 10
afternoon <- rep(0:1,N_visits*N_cafes/2)
cafe_id <- rep( 1:N_cafes , each=N_visits )
mu <- a_cafe[cafe_id] + b_cafe[cafe_id]*afternoon
sigma <- 0.5  # std dev within cafes
wait <- rnorm( N_visits*N_cafes , mu , sigma )
d <- data.frame( cafe=cafe_id , afternoon=afternoon , wait=wait )
```

# 14.1.3. The varying slopes model.
```{r}
R <- rlkjcorr( 1e4 , K=2 , eta=2 )
dens( R[,1,2] , xlab="correlation" )
```

```{r}
m14.1 <- ulam(
    alist(
        wait ~ normal( mu , sigma ),
        mu <- a_cafe[cafe] + b_cafe[cafe]*afternoon,
        c(a_cafe,b_cafe)[cafe] ~ multi_normal( c(a,b) , Rho , sigma_cafe ),
        a ~ normal(5,2),
        b ~ normal(-1,0.5),
        sigma_cafe ~ exponential(1),
        sigma ~ exponential(1),
        Rho ~ lkj_corr(2)
    ) , data=d , chains=4 , cores=2,iter=2000,log_lik=TRUE)
# What is lkj_corr?
?rlkjcorr
?rmultinom # for multi_ormal, correct?
precis(m14.1,depth=3)

```
# the end of varying slope test

# problems
## 13M3. Re-estimate the varying slopes model for the UCBadmit data, now using a non-centered parameterization. Compare the efficiency of the forms of the model, using n_eff. Which is better? Which chain sampled faster?

* modified original m11.8 quad version into ulam
```{r}
data(UCBadmit)
d <- UCBadmit
dat <- list( A=d$admit , N=d$applications , gid=ifelse( d$applicant.gender=="male" , 1 , 2 ),did=rep(1:6,each=2))
dat
m11.8.ulam <- ulam(
    alist(
        A ~ dbinom(N, p),
        logit(p) <- a[gid] + delta[did] ,
        a[gid] ~ dnorm( 0 , 1.5 ) ,
        delta[did] ~ dnorm( 0 , 1.5 )
    ) , data=dat,chains=4,cores=2,iter=4000,log_lik=TRUE)
precis( m11.8.ulam , depth=2 )
plot(m11.8.ulam , depth=2)
```

* multitire? partial pooling model? 
```{r}
m.practice13M3.a <- ulam(
    alist(
        A ~ dbinom(N , p ) ,
        ## partial pooling model?
        logit(p) <- a[did] + delta[did]*gid,
          a[did] ~ dnorm( 0 , 1.5 ) ,
          delta[did] ~ dnorm(0,1.5) 
        ) , data=dat , chains=4 , cores=2 , iter=4000,log_lik=TRUE )
precis(m.practice13M3.a, depth=2)
plot(m.practice13M3.a,depth=2)
```

* the varying slopes model. a_did and b_did are related (correct wording?)
```{r}
m.practice13M3.b <- ulam(
    alist(
        A ~ dbinom(N,p),
        p <- a_did[did] + b_did[did]*gid,
        c(a_did,b_did)[did] ~ multi_normal( c(a,b) , Rho , sigma_did ),
        a ~ normal(0,1.5),
        b ~ normal(0,1.5),
        sigma_did ~ exponential(1),
        Rho ~ lkj_corr(2)
    ) , data=dat , chains=4 , cores=2 , iter=4000,log_lik=TRUE)
# error... why?
precis(m.practice13M3.b, depth=2)
plot(m.practice13M3.b,depth=2)
```
# normal? dnorm?
```{r}
m.practice13M3.c <- ulam(
    alist(
        A ~ dbinom(N,p),
        p <- a_did[did] + b_did[did]*gid,
        c(a_did,b_did)[did] ~ multi_normal( c(a,b) , Rho , sigma_did ),
        a ~ dnorm(0,1.5),
        b ~ dnorm(0,1.5),
        sigma_did ~ exponential(1),
        Rho ~ lkj_corr(2)
    ) , data=dat , chains=4 , cores=2 , iter=4000,log_lik=TRUE)
# error why?
precis(m.practice13M3.c, depth=2)
plot(m.practice13M3.c,depth=2)
```

* compare
```{r}
compare(m11.8.ulam,m.practice13M3.a,m.practice13M3.b,m.practice13M3.c)
```


# sessionInfo()
```{r}
sessionInfo()
```

