---
title: "Chap11_Kazu"
author: "Kazu"
date: "11/1/2019"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=TRUE,cache = FALSE)
library(rethinking)
library(lme4);library(lmerTest);library(ggplot2);library(reshape2);library(tidyverse);library(readr)
```

# 
```{r}
data(chimpanzees)
d <- chimpanzees
```
* There are four combinations:
* (1) prosoc_left= 0 and condition= 0: Two food items on right and no partner. 
* (2) prosoc_left= 1 and condition= 0: Two food items on left and no partner.
* (3) prosoc_left= 0 and condition= 1: Two food items on right and partner present. 
* (4) prosoc_left= 1 and condition= 1: Two food items on left and partner present.

* let’s build an index variable containing the values 1 through 4, to index the combinations above.
```{r R code 11.2}
d$treatment <- 1 + d$prosoc_left + 2*d$condition
```

* Now treatment contains the values 1 through 4, matching the numbers in the list above. You can verify by using cross-tabs
```{r R code 11.3}
 xtabs( ~ treatment + prosoc_left + condition , d )
```
```{r R code 11.4}
m11.1 <- quap(
    alist(
        pulled_left ~ dbinom( 1 , p ) ,
        logit(p) <- a ,
        a ~ dnorm( 0 , 10 )
) , data=d )
```

# 
```{r R code 11.5}
set.seed(1999)
prior <- extract.prior( m11.1 , n=1e4 )
```

# 
```{r R code 11.6}
p <- inv_logit( prior$a )
dens( p , adj=0.1 )
```

# 
```{r R code 11.7}
m11.2 <- quap(
    alist(
        pulled_left ~ dbinom( 1 , p ) ,
        logit(p) <- a + b[treatment] ,
        a ~ dnorm( 0 , 1.5 ),
        b[treatment] ~ dnorm( 0 , 10 )
) , data=d )
set.seed(1999)
prior <- extract.prior( m11.2 , n=1e4 )
p <- sapply( 1:4 , function(k) inv_logit( prior$a + prior$b[,k] ) )
```

# 
```{r R code 11.8}
 dens( abs( p[,1] - p[,2] ) , adj=0.1 )
```

# 
```{r R code 11.9}
m11.3 <- quap(
    alist(
        pulled_left ~ dbinom( 1 , p ) ,
        logit(p) <- a + b[treatment] ,
        a ~ dnorm( 0 , 1.5 ),
        b[treatment] ~ dnorm( 0 , 0.5 )
    ) , data=d )
set.seed(1999)
prior <- extract.prior( m11.3 , n=1e4 )
p <- sapply( 1:4 , function(k) inv_logit( prior$a + prior$b[,k] ) )
mean( abs( p[,1] - p[,2] ) )
```

#
```{r}
# prior trimmed data list
dat_list <- list(
    pulled_left = d$pulled_left,
    actor = d$actor,
    treatment = as.integer(d$treatment) )
# particles in 11-dimensional space
m11.4 <- ulam(
    alist(
        pulled_left ~ dbinom( 1 , p ) ,
        logit(p) <- a[actor] + b[treatment] ,
        a[actor] ~ dnorm( 0 , 1.5 ),
        b[treatment] ~ dnorm( 0 , 0.5 )
),
    data=dat_list , chains=4 )
precis( m11.4 , depth=2 )
```

#
```{r}
post <- extract.samples(m11.4)
p_left <- inv_logit( post$a )
plot( precis( as.data.frame(p_left) ) , xlim=c(0,1) )
```

# 
```{r R code 11.12}
labs <- c("R/N","L/N","R/P","L/P")
plot( precis( m11.4 , depth=2 , pars="b" ) , labels=labs )
```

# 
```{r R code 11.13}
diffs <- list(
    db13 = post$b[,1] - post$b[,3],
    db24 = post$b[,2] - post$b[,4] )
plot( precis(diffs) )
```

# 
```{r R code 11.14}
pl <- by( d$pulled_left , list( d$actor , d$treatment ) , mean )
pl[1,]
```

#
```{r R code 11.15}
plot( NULL , xlim=c(1,28) , ylim=c(0,1) , xlab="" ,
    ylab="proportion left lever" , xaxt="n" , yaxt="n" )
axis( 2 , at=c(0,0.5,1) , labels=c(0,0.5,1) )
abline( h=0.5 , lty=2 )
for ( j in 1:7 ) abline( v=(j-1)*4+4.5 , lwd=0.5 )
for ( j in 1:7 ) text( (j-1)*4+2.5 , 1.1 , concat("actor ",j) , xpd=TRUE )
for ( j in (1:7)[-2] ) {
    lines( (j-1)*4+c(1,3) , pl[j,c(1,3)] , lwd=2 , col=rangi2 )
    lines( (j-1)*4+c(2,4) , pl[j,c(2,4)] , lwd=2 , col=rangi2 )
}
points( 1:28 , t(pl) , pch=16 , col="white" , cex=1.7 )
points( 1:28 , t(pl) , pch=c(1,1,16,16) , col=rangi2 , lwd=2 )
yoff <- 0.01
text( 1 , pl[1,1]-yoff , "R/N" , pos=1 , cex=0.8 )
text( 2 , pl[1,2]+yoff , "L/N" , pos=3 , cex=0.8 )
text( 3 , pl[1,3]-yoff , "R/P" , pos=1 , cex=0.8 )
text( 4 , pl[1,4]+yoff , "L/P" , pos=3 , cex=0.8 )
mtext( "observed proportions\n" )
```

# 
```{r R code 11.16}
dat <- list( actor=rep(1:7,each=4) , treatment=rep(1:4,times=7) )
p_post <- link_ulam( m11.4 , data=dat )
p_mu <- apply( p_post , 2 , mean )
p_ci <- apply( p_post , 2 , PI )
```

# 
```{r R code 11.17}
d$side <- d$prosoc_left + 1 # right 1, left 2
d$cond <- d$condition + 1 # no partner 1, partner 
```

#
```{r R code 11.18}
dat_list2 <- list(
    pulled_left = d$pulled_left,
    actor = d$actor,
    side = d$side,
    cond = d$cond )
m11.5 <- ulam(
    alist(
        pulled_left ~ dbinom( 1 , p ) ,
        logit(p) <- a[actor] + bs[side] + bc[cond] ,
        a[actor] ~ dnorm( 0 , 1.5 ),
        bs[side] ~ dnorm( 0 , 0.5 ),
        bc[cond] ~ dnorm( 0 , 0.5 )
),
data=dat_list2 , chains=4 , log_lik=TRUE )
```

#
```{r R code 11.19}
compare( m11.5 , m11.4 , func=LOO )
```

# 
```{r R code 11.20}
 post <- extract.samples( m11.4 , clean=FALSE )
str(post)
```

# 
```{r R code 11.21}
m11.4_stan_code <- stancode(m11.4)
m11.4_stan <- stan( model_code=m11.4_stan_code , data=dat_list , chains=4 )
compare( m11.4_stan , m11.4 )
```

# 11.1.2. Relative shark and absolute penguin. 
```{r R code 11.22}
post <- extract.samples(m11.4)
mean( exp(post$b[,4]-post$b[,2]) )
```

# 11.1.3. Aggregated binomial: Chimpanzees again, condensed.
```{r}
data(chimpanzees)
d <- chimpanzees
d$treatment <- 1 + d$prosoc_left + 2*d$condition
d$side <- d$prosoc_left + 1 # right 1, left 2
d$cond <- d$condition + 1 # no partner 1, partner 2
d_aggregated <- aggregate(
    d$pulled_left ,
    list( treatment=d$treatment , actor=d$actor ,
          side=d$side , cond=d$cond ) ,
    sum )
colnames(d_aggregated)[5] <- "left_pulls"
```

#
```{r R code 11.24}
dat <- with( d_aggregated , list(
    left_pulls = left_pulls,
    treatment = treatment,
    actor = actor,
    side = side,
    cond = cond ) )
m11.6 <- ulam(
    alist(
        left_pulls ~ dbinom( 18 , p ) ,
        logit(p) <- a[actor] + b[treatment] ,
        a[actor] ~ dnorm( 0 , 1.5 ) ,
        b[treatment] ~ dnorm( 0 , 0.5 )
),
data=dat , chains=4 , log_lik=TRUE )
```

```{r R code 11.25}
compare( m11.6 , m11.4 , func=LOO )
```

```{r}
# deviance of aggregated 6-in-9
-2*dbinom(6,9,0.2,log=TRUE)
# deviance of dis-aggregated
-2*sum(dbern(c(1,1,1,1,1,1,0,0,0),0.2,log=TRUE))
```

```{r R code 11.27}
( k <- LOOPk(m11.6) )
```

# 11.1.4. Aggregatedbinomial:Graduateschooladmissions.
```{r R code 11.28}
data(UCBadmit)
d <- UCBadmit
```

```{r R code 11.29}
d$gid <- ifelse( d$applicant.gender=="male" , 1 , 2 )
m11.7 <- quap(
    alist(
        admit ~ dbinom( applications , p ) ,
        logit(p) <- a[gid] ,
        a[gid] ~ dnorm( 0 , 1.5 )
    ) , data=d )
precis( m11.7 , depth=2 )
```

```{r R code 11.30}
post <- extract.samples(m11.7)
diff_a <- post$a[,1] - post$a[,2]
diff_p <- inv_logit(post$a[,1]) - inv_logit(post$a[,2])
precis( list( diff_a=diff_a , diff_p=diff_p ) )
```

```{r R code 11.31}
postcheck( m11.7 , n=1e4 )
# draw lines connecting points from same dept
d$dept_id <- rep( 1:7 , each=2 )
for ( i in 1:6 ) {
    x <- 1 + 2*(i-1)
    y1 <- d$admit[x]/d$applications[x]
    y2 <- d$admit[x+1]/d$applications[x+1]
    lines( c(x,x+1) , c(y1,y2) , col=rangi2 , lwd=2 )
    text( x+0.5 , (y1+y2)/2 + 0.05 , d$dept[x] , cex=0.8 , col=rangi2 )
}
```

```{r R code 11.32}
d$dept_id <- rep(1:6,each=2)
m11.8 <- quap(
    alist(
        admit ~ dbinom( applications , p ) ,
        logit(p) <- a[gid] + delta[dept_id] ,
        a[gid] ~ dnorm( 0 , 1.5 ) ,
        delta[dept_id] ~ dnorm( 0 , 1.5 )
    ) , data=d )
precis( m11.8 , depth=2 )
```

```{r R code 11.33}
post <- extract.samples(m11.8)
diff_a <- post$a[,1] - post$a[,2]
diff_p <- inv_logit(post$a[,1]) - inv_logit(post$a[,2])
precis( list( diff_a=diff_a , diff_p=diff_p ) )
```

```{r R code 11.34}
pg <- sapply( 1:6 , function(k)
    d$applications[d$dept_id==k]/sum(d$applications[d$dept_id==k]) )
rownames(pg) <- c("male","female")
colnames(pg) <- unique(d$dept)
round( pg , 2 )
```

# 11.1.5. Multinomialandcategoricalmodels.  (optional)

# 11.5 Practice
* 10E1. If an event has probability 0.35, what are the log-odds of this event?
```{r}
log(0.35/(1-0.35))
logit(0.35)
```

* 10E2. If an event has log-odds 3.2, what is the probability of this event?
```{r}
inv_logit(3.2) # 0.9608343
logit(inv_logit(3.2)) # 3.2
```


* 10E3. Suppose that a coefficient in a logistic regression has value 1.7. What does this imply about the proportional change in odds of the outcome?
```{r}

```
* 10M1. As explained in the chapter,binomial data can be organized in aggregated and disaggregated forms, without any impact on inference. But the likelihood of the data does change when the data are converted between the two formats. Can you explain why?
```{r}

```

