---
title: "Chap13_Jan31_2020_HW_Kazu"
author: "Kazu"
date: "1/31/2020"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

# 13. Models with memory
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=TRUE,cache = FALSE)
library(rethinking)
library(lme4);library(lmerTest);library(ggplot2);library(reshape2);library(tidyverse);library(readr)
```

Please find attached csv.  This csv has measurements of tomato internodes and petioles from wildtype (Moneymaker) and various phytochrome mutant lines.  Measurements were made at 3 time points, 21, 28, and 35 days after germination under two treatments, simulated sun and simulated shade.

for today let's focus on day 35.  Also let's focus on total stem length.  So: first

# Q1)
## a) subset the data for day 35
```{r}
F4phyE<-read_csv("../figure4phyE.csv")
summary(F4phyE)
F4phyE.day35 <- F4phyE %>% filter(day==35)
summary(F4phyE.day35)
```

b) create a new column "stem_length" that is the sum of epi, int1, int2, and int3
```{r}
F4phyE.day35.v2 <- F4phyE.day35 %>% mutate(stem_length=epi+int1+int2+int3) # %>% View()
F4phyE.day35.v2
```

c) although flats are listed as 1-6, flats in sun and shade are separate. Create a new column "flat2" that corrects for this.
```{r}
F4phyE.day35.v3 <- F4phyE.day35.v2 %>% unite("flat2",c("treatment","flat"),sep="_", remove=FALSE)
F4phyE.day35.v3
```


Ultimately you want to know if any of the mutants have a different length from Moneymaker, in sun or in shade, or if the response to shade differs.
```{r}
# sun treatment is zero or one
dat_list.Q1.zero.one<- list(
    stem_length = F4phyE.day35.v3$stem_length,
    genotype = as.integer(factor(F4phyE.day35.v3$genotype,levels=c("Moneymaker", 
                                            "phyB1", 
                                            "phyB2", 
                                            "phyB1/B2", 
                                            "phyEami3", 
                                            "phyEami7"))),
    flat_id = as.integer(factor(F4phyE.day35.v3$flat2,levels=c("shade_1",
                                                            "shade_2",
                                                            "shade_3",
                                                            "shade_4",
                                                            "shade_5",
                                                            "shade_6",
                                                            "sun_1",
                                                            "sun_2",
                                                            "sun_3",
                                                            "sun_4",
                                                            "sun_5",
                                                            "sun_6"))),
    treatment = ifelse(F4phyE.day35.v3$treatment=="sun",0,1)) 
# sun treatment is 1 or 2.
dat_list.Q1.one.two<- list(
    stem_length = F4phyE.day35.v3$stem_length,
    genotype = as.integer(factor(F4phyE.day35.v3$genotype,levels=c("Moneymaker", 
                                            "phyB1", 
                                            "phyB2", 
                                            "phyB1/B2", 
                                            "phyEami3", 
                                            "phyEami7"))),
    flat_id = as.integer(factor(F4phyE.day35.v3$flat2,levels=c("shade_1",
                                                            "shade_2",
                                                            "shade_3",
                                                            "shade_4",
                                                            "shade_5",
                                                            "shade_6",
                                                            "sun_1",
                                                            "sun_2",
                                                            "sun_3",
                                                            "sun_4",
                                                            "sun_5",
                                                            "sun_6"))),
    treatment = ifelse(F4phyE.day35.v3$treatment=="sun",1,2)) 
# "sun" is 1, "shade" is 2
dat_list.Q1.zero.one

# 
str(ifelse(F4phyE.day35.v3$treatment=="sun",0,1)) # numeric
str(ifelse(F4phyE.day35.v3$treatment=="sun",0L,1L)) # integer

```

# look the raw data
```{r}
str(F4phyE.day35.v3)
F4phyE.day35.v3$treatment <-factor(F4phyE.day35.v3$treatment,levels=c("sun","shade"))
F4phyE.day35.v3 %>% View()
F4phyE.day35.v3 %>% ggplot(aes(x=treatment, y=stem_length,color=treatment)) + geom_jitter() + facet_grid(.~genotype)
# sample numbers (pivot table, tidyverse way)
F4phyE.day35.v3 %>% group_by(genotype,treatment,flat) %>% summarize(num=n()) %>% spread(flat,num)
# incomplete block design
with(F4phyE.day35.v3, table(genotype,flat,treatment)) # traditional way, creating list

```

* Looks there are treatment effects are genotype effects.

# Q2) Fit 3 models, all of which include genotype and treatment
* Data format depends on indexing or not.

Q2: a) do not include flat

## ModelQ1.a.wrong
* Kazu'Q: When\ logit\ should\ be\ used? \\
* Use logit when value is between 0 and 1, likely probability. (correct?)
* So this is wrong

$$
stem_length_i \sim Normal(0,p) \\
logit(p) = \alpha_{genotype} + \beta_{treatment} \\
\alpha_{group} \sim Normal(0, 1) \\
\beta_{treatment} \sim Normal(0, 1)
$$

# wrong model with logit
```{r}
set.seed(15)
# error
mQ2.a.zero.one <- ulam(
    alist(
        stem_length ~ dnorm( 1 , p),
        logit(p) <- a[genotype] + b[treatment],
        b[treatment] ~ dnorm( 0 , 1),
        a[genotype] ~ dnorm(0 , 1)
    ), data=dat_list.Q1.zero.one , chains=2 , cores=2 , log_lik=TRUE )
# no error, why?
mQ1.a.one.two <- ulam(
    alist(
        stem_length ~ dnorm( 1 , p),
        logit(p) <- a[genotype] + b[treatment],
        b[treatment] ~ dnorm( 0 , 1),
        a[genotype] ~ dnorm(0 , 1)
    ), data=dat_list.Q1,one.two , chains=2 , cores=2 , log_lik=TRUE )
```

# Use normal distribution with mu and sigma also intercept
## # mQ1.a2  intercept two treatment (no interaction)
* Is this correct formula?

$$
stem\_length \sim Normal(mu,sigma) \\
mu_{genotype},_{treatment} = \alpha_{genotype} + \beta\_g_{genotype}*genotype + \beta\_t_{treatment}*treatment \\
\alpha_{genotype} \sim Normal(140,50) , for\ genotype\\
\beta\_g_{genotype} \sim Normal(0, 1) , for\ genotype\\
\beta\_t_{treatment} \sim Normal(0, 1) , for\ treatment \\
\sigma \sim Exponential(0,1)
$$

```{r}
# estimate intercept using mean?. How to translate into ulam() format for alpha genotype? Using index is easy?
F4phyE.day35.v3 %>% summarize(mean(stem_length),sd(stem_length)) # 141 48.5
# this use common alpha
mQ2.a2.nonindex.zero.one <- ulam(
    alist(
        stem_length ~ dnorm(mu , sigma),
        mu <- alpha + b_g*genotype + b_t*treatment,
        alpha ~ dnorm(140,50),
        b_g ~ dnorm( 0 , 1),
        b_t ~ dnorm(0 , 1),
        sigma ~ dexp( 1 )
    ), data=dat_list.Q1.zero.one , chains=2 , cores=2 , log_lik=TRUE )

precis(mQ1.a2.nonindex.zero.one,depth=2)
pairs(mQ1.a2.nonindex.zero.one)

```

```{r}
mQ2.a2.index.one.two <- ulam(
    alist(
        stem_length ~ dnorm(mu , sigma),
        mu <- alpha[genotype] + b_g*[genotype] + b_t*treatment,
        alpha[genotype] ~ dnorm(140,50),
        b_g~ dnorm( 0 , 1),
        b_t ~ dnorm(0 , 1),
        sigma ~ dexp( 1 )
    ), data=dat_list.Q1.one.two , chains=2 , cores=2 , log_lik=TRUE )

precis(mQ1.a2.nonindex.one.two,depth=2)

```

# 

$$
stem\_length \sim Normal(mu,sigma) \\
mu_{genotype},_{treatment} = \alpha + \beta\_g_{genotype}*genotype + \beta\_t_{treatment}*treatment \\
\alpha \sim Normal(140,50) , common\\
\beta\_g_{genotype} \sim Normal(0, 1) , for\ genotype\\
\beta\_t_{treatment} \sim Normal(0, 1) , for\ treatment \\
\sigma \sim Exponential(0,1)
$$
# 
```{r}
mQ2.a2.index.zero.one <- ulam(
    alist(
        stem_length ~ dnorm(mu , sigma),
        mu <- alpha + b_g[genotype] + b_t*treatment,
        alpha ~ dnorm(140,50),
        b_g[genotype] ~ dnorm( 0 , 1),
        b_t ~ dnorm(0 , 1),
        sigma ~ dexp( 1 )
    ), data=dat_list.Q1.zero.one , chains=2 , cores=2 , log_lik=TRUE )

precis(mQ1.a2.index.zero.one,depth=2)
compare(mQ1.a2.zero.one,mQ1.a2.index.zero.one)
```

```{r}
mQ2.a2.index.one.two <- ulam(
    alist(
        stem_length ~ dnorm(mu , sigma),
        mu <- alpha + b_g[genotype] + b_t*treatment,
        alpha ~ dnorm(140,50),
        b_g[genotype] ~ dnorm( 0 , 1),
        b_t ~ dnorm(0 , 1),
        sigma ~ dexp( 1 )
    ), data=dat_list.Q1.one.two , chains=2 , cores=2 , log_lik=TRUE )

precis(mQ1.a2.index.one.two,depth=2)
```

```{r}
mQ2.a2.trtindex.zero.one <- ulam(
    alist(
        stem_length ~ dnorm(mu , sigma),
        mu <- alpha + b_g*genotype + b_t[treatment],
        alpha ~ dnorm(140,50),
        b_g ~ dnorm( 0 , 1),
        b_t[treatment] ~ dnorm(0 , 1),
        sigma ~ dexp( 1 )
    ), data=dat_list.Q1.zero.one , chains=2 , cores=2 , log_lik=TRUE )
# error. "indexing" is not copatible with "0".
precis(mQ1.a2.trtindex.zero.one,depth=2)
```

```{r}
mQ2.a2.trtindex.one.two <- ulam(
    alist(
        stem_length ~ dnorm(mu , sigma),
        mu <- alpha + b_g*genotype + b_t[treatment],
        alpha ~ dnorm(140,50),
        b_g ~ dnorm( 0 , 1),
        b_t[treatment] ~ dnorm(0 , 1),
        sigma ~ dexp( 1 )
    ), data=dat_list.Q1.one.two , chains=2 , cores=2 , log_lik=TRUE )

precis(mQ1.a2.trtindex.one.two,depth=2)
```

* indexing gt as well as treatment
```{r}
mQ1.a2.trtgtindex.one.two <- ulam(
    alist(
        stem_length ~ dnorm(mu , sigma),
        mu <- alpha + b_g[genotype] + b_t[treatment],
        alpha ~ dnorm(140,50),
        b_g[genotype] ~ dnorm( 0 , 1),
        b_t[treatment] ~ dnorm(0 , 1),
        sigma ~ dexp( 1 )
    ), data=dat_list.Q1.one.two , chains=2 , cores=2 , log_lik=TRUE )

precis(mQ1.a2.trtgtindex.one.two,depth=2)
```



# how about to no intercept?
## # mQ1.a3  no intercept, two treatment (no interaction)

$$
stem\_length \sim Normal(mu,sigma) \\
mu =   \beta\_g_{genotype} + \beta\_t_{treatment} \\
\alpha \sim Normal(140,50) \\
\beta\_g \sim Normal(0, 1) \\
\beta\_t_ \sim Normal(0, 1) \\
\sigma \sim Exponential(0,1)
$$
```{r}
mQ2.a3.nonindex.zero.one <- ulam(
    alist(
        stem_length ~ dnorm(mu , sigma),
        mu <- b_g*genotype + b_t*treatment,
        b_g ~ dnorm( 0 , 1),
        b_t ~ dnorm(0 , 1),
        sigma ~ dexp( 1 )
    ), data=dat_list.Q1.zero.one , chains=2 , cores=2 , log_lik=TRUE )

precis(mQ1.a3.nonindex.zero.one,depth=2)
pairs(mQ1.a3.nonindex.zero.one)
compare(mQ1.a2.nonindex.zero.one,mQ1.a3.nonindex.zero.one)
# no intercept is worse.
```

# interaction model
* Is this correct model description? with no index way?

$$
stem\_length \sim Normal(mu,sigma) \\
mu =  \alpha +  \beta\_g_{genotype} + \beta\_t_{treatment} + \beta\_gt_{genotype*treatment} \\
\alpha \sim Normal(140,50) \\
\beta\_g \sim Normal(0, 1) \\
\beta\_t_ \sim Normal(0, 1) \\
\beta\_gt_ \sim Normal(0, 1) \\
\sigma \sim Exponential(0,1)
$$
```{r}
mQ1.a4.nonindex.zero.one <- ulam(
    alist(
        stem_length ~ dnorm(mu , sigma),
        mu <- alpha + b_g*genotype + b_t*treatment + b_gt*genotype*treatment,
        alpha ~ dnorm(140,50),
        b_g ~ dnorm( 0 , 1),
        b_t ~ dnorm(0 , 1),
        b_gt ~ dnorm(0, 1),
        sigma ~ dexp( 1 )
    ), data=dat_list.Q1.zero.one , chains=2 , cores=2 , log_lik=TRUE )

precis(mQ1.a4.nonindex.zero.one,depth=2)
compare(mQ1.a2.nonindex.zero.one,mQ1.a4.nonindex.zero.one)
# little improvement
pairs(mQ1.a4.nonindex.zero.one) # I could see negative correlation between alpha and b_g, but not b_gt.
```


# using index method for interaction model (only genotype is indexed)
$$
stem\_length \sim Normal(mu,sigma) \\
mu =  \alpha_{genotype} +  \beta\_g_{genotype}*genotype + \beta\_t_{treatment}*treatment + \beta\_gt_{genotype*treatment}*genotype*treatment \\
\alpha \sim Normal(140,50) \\
\beta\_g \sim Normal(0, 1) \\
\beta\_t_ \sim Normal(0, 1) \\
\beta\_gt_ \sim Normal(0, 1) \\
\sigma \sim Exponential(0,1)
$$



```{r}
mQ1.a4.gtindex.one.two <- ulam(
    alist(
        stem_length ~ dnorm(mu , sigma),
        mu <- alpha + b_g[genotype] + b_t*treatment + b_gt[genotype]*treatment,
        alpha ~ dnorm(140,50),
        b_g[genotype] ~ dnorm( 0 , 1),
        b_t ~ dnorm(0 , 1),
        b_gt[genotype] ~ dnorm (0,1),
        sigma ~ dexp( 1 )
    ), data=dat_list.Q1.one.two , chains=2 , cores=2 , log_lik=TRUE )
# Julin's priors
mQ2.a4.gtindex.one.two.2 <- ulam(
    alist(
        stem_length ~ dnorm(mu , sigma),
        mu <- alpha + b_g[genotype] + b_t*treatment + b_gt[genotype]*treatment,
        alpha ~ dnorm(125,50),
        b_g[genotype] ~ dnorm( 0 , 50),
        b_t ~ dnorm(0 , 50),
        b_gt[genotype] ~ dnorm (0,50),
        sigma ~ dexp( 1 )
    ), data=dat_list.Q1.one.two , chains=2 , cores=2 , log_lik=TRUE )

precis(mQ1.a4.gtindex.one.two.2,depth=2)
compare(mQ1.a4.gtindex.one.two,mQ1.a4.gtindex.one.two.2) 
pairs(mQ1.a4.gtindex.one.two.2) # different priors. Julin's priors gave me strong  negative correlation betweeen b_t and b_gts!!! And Julin's prior is better model!
compare(mQ1.a4.gtindex.one.two.2,mQ1.a4.nonindex.zero.one.2) # gt index is better than nonindex.

```

# using index method for interaction model (only genotype is indexed). no intercept
```{r}
mQ2.a4.gtindex.one.two.nointercept <- ulam(
    alist(
        stem_length ~ dnorm(mu , sigma),
        mu <- b_g[genotype] + b_t*treatment + b_gt[genotype]*treatment,
        #alpha ~ dnorm(140,50),
        b_g[genotype] ~ dnorm( 140 , 50),
        b_t ~ dnorm(0 , 1),
        b_gt[genotype] ~ dnorm (0,1),
        sigma ~ dexp( 1 )
    ), data=dat_list.Q1.one.two , chains=2 , cores=2 , log_lik=TRUE )

compare(mQ1.a4.gtindex.one.two.2,mQ1.a4.nonindex.zero.one.2) # gt index is better than nonindex.

```



Q3) Compare the models, which is preferred?
# model comparison
```{r}
compare(mQ2.a2.nonindex.zero.one,
        mQ2.a2.index.zero.one,
        mQ2.a2.nonindex.one.two,
        mQ2.a2.index.one.two,
        mQ2.a2.trtindex.one.two,
        mQ2.a2.trtgtindex.one.two)
```
* All of the same.









# Q to Julin: How did you chose your priors?


######### under construction ##########
Q2: b) include flat without pooling
* I feel I do understand what pooling is meant....
* Model description

$$
stem\_length \sim Normal(mu,sigma) \\
mu =   \beta\_g_{genotype} + \beta\_t_{treatment} + \beta\_f_{flat} \\
\beta\_g \sim Normal(140,50) \\
\beta\_t \sim Normal(0, 1) \\
\beta\_f \sim Normal(0,1) \\
\sigma \sim Exponential(0,1)
$$



```{r eval=FALSE}
set.seed(15)
mQ4.b1.gtindex.one.two <- ulam(
    alist(
        stem_length ~ dnorm(mu , sigma),
        mu <- alpha + b_g[genotype] + b_t*treatment + b_gt[genotype]*treatment,
        alpha ~ dnorm(125,50),
        b_g[genotype] ~ dnorm( 0 , 50),
        b_t ~ dnorm(0 , 50),
        b_gt[genotype] ~ dnorm (0,50),
        sigma ~ dexp( 1 )
    ), data=dat_list.Q1.one.two , chains=2 , cores=2 , log_lik=TRUE )

```


Q2: c) use a hierarchical model that allows partial pooling across flats

$$
stem\_length[i] \sim Normal(mu,sigma) \\
mu[i] =   \beta\_g_{genotype[i]} + \beta\_t_{treatment[i]} \\
\alpha \sim Normal(140,50) \\
\beta\_g \sim Normal(0, 1) \\
\beta\_t \sim Normal(0, 1) \\
\sigma \sim Exponential(0,1)
$$

```{r eval=FALSE}

```


# Q3) Compare the models, which is preferred?
```{r eval=FALSE}
compare(mQ1.a,mQ1.b,mQ1.c)
```

######## work with Julin's model and come back above later ########
```{r}
data <- read_csv(file.path("..","figure4phyE.csv"))
data35 <- data %>%
  filter(day==35) %>%
  mutate(stem_length=epi + int1 + int2 + int3,
         flat2=as.integer(as.factor(str_c(treatment, flat))),
         shade_i=ifelse(treatment=="sun", 0L, 1L),
         g_i= as.integer(factor(genotype,
                                   levels=c("Moneymaker", 
                                            "phyB1", 
                                            "phyB2", 
                                            "phyB1/B2", 
                                            "phyEami3", 
                                            "phyEami7")))) %>%
  select(genotype, treatment, g_i, shade_i, flat2, stem_length)
data35
datsmall <- data35 %>% select(stem_length, shade_i, g_i, flat2) # Julin's version
# from my data
#datsmall<-dat_list.Q1.zero.one %>%  bind_cols() %>% rename(shade_i=treatment, g_i=genotype,flat2=flat_id) 
  # convert mine to Julin's version (not necessary to be list object)

mq2c2 <- ulam(alist(stem_length ~ dnorm(mu,sigma),
                  mu <- alpha[g_i] + b_shade[g_i]*shade_i + b_fl[flat2],
                  alpha[g_i] ~ dnorm(125,50),
                  b_shade[g_i] ~ dnorm(0, 50),
                  b_fl[flat2] ~ dnorm(0,sigma_fl),
                  sigma ~ dexp(1),
                  sigma_fl ~ dcauchy(0,3)),
            data=datsmall,
            chains=4,
            cores=2,
            iter=4000,
            log_lik = TRUE)
precis(mq2c2, depth=2)
traceplot(mq2c2, ask=FALSE)
pairs(mq2c2) # Error in plot.new() : figure margins too large
# alternate of pairs (by Julin)
extract.samples(mq2c2) %>%
  as.data.frame() %>%
  cor() %>%
  round(2)

```

# Q4) Using the hierarchical model, make posterior predictions
a) for average cluster
```{r}
post <- extract.samples(mq2c2) # n=1000 is default, but why 8000? coming fomr iter = 4000 in the model?
names(post)
str(post)
```
# predicted value function? (Julin)
```{r}
link_avg <- function(genotype, shade) {
  with(post, alpha[,genotype] + shade*b_shade[,genotype])
}
```
# create a data frame to hold the results (Julin)
```{r}
pred.df <- data35 %>% 
  select(-stem_length, -flat2) %>% 
  unique() %>%
  mutate(treatment=factor(treatment, levels = c("sun", "shade")) ) # so the plot order is correct
pred.df
```

# summary table of predicted value
```{r}
pred.df.avg <- pred.df %>%
  mutate(average_response=map2(g_i, shade_i, link_avg),
         pet_length=map_dbl(average_response, mean),
         low.89=map_dbl(average_response, ~ HPDI(.)[1]),
         high.89=map_dbl(average_response, ~ HPDI(.)[2]))
pred.df.avg
```

# plot it
```{r}
pred.df.avg %>%
  ggplot(aes(x=genotype, y=pet_length, ymin=low.89, ymax=high.89, fill=treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(width = .5, position = position_dodge(width=.9)) +
  ggtitle("prediction averaged across flat")
```

b) for same clusters
```{r}

```

c) showing the "marginal" from cluster
```{r}

```

d) showing new clusters.
```{r}

```

Q5) Reparameterize the model to help with divergent transitions (even if there aren't any)
```{r}

```

Q6--optional)
a) Which genotypes differ from MoneyMaker in Sun conditions?
* (J's) We compare posterior estimates for alpha for Moneymaker vs the others and ask how often the difference is non-zero in a particular direction
```{r}
avg.MM.sun <- pred.df.avg %>%
  filter(genotype=="Moneymaker", treatment=="sun") %>%
  pull(average_response) %>% unlist()

#can't get this to work with Map, unfortunately...
post.MM.sun.contrasts <- pred.df.avg %>%
  filter(genotype!="Moneymaker", treatment=="sun") %>%
  pull(average_response) %>%
  sapply(function(x) unlist(x) - avg.MM.sun) %>%
  as.data.frame()

colnames(post.MM.sun.contrasts) <- pred.df.avg  %>%
  filter(genotype!="Moneymaker", treatment=="sun") %>%
  pull(genotype)

head(post.MM.sun.contrasts)
precis(post.MM.shade.contrasts)
plot(precis(post.MM.shade.contrasts))

```

b) Which genotypes differ from MoneyMaker in Shade conditions?
```{r}
avg.MM.shade <- pred.df.avg %>%
  filter(genotype=="Moneymaker", treatment=="shade") %>%
  pull(average_response) %>% unlist()

#can't get this to work with Map, unfortunately...
post.MM.shade.contrasts <- pred.df.avg %>%
  filter(genotype!="Moneymaker", treatment=="shade") %>%
  pull(average_response) %>%
  sapply(function(x) unlist(x) - avg.MM.shade) %>%
  as.data.frame()

colnames(post.MM.shade.contrasts) <- pred.df.avg  %>%
  filter(genotype!="Moneymaker", treatment=="shade") %>%
  pull(genotype)

head(post.MM.shade.contrasts)
precis(post.MM.shade.contrasts)
plot(precis(post.MM.shade.contrasts))
```

c) Which genotypes differ from MoneyMaker in their response to shade (difference in sun vs shade)?
```{r}
plot(precis(post.MM.shade.contrasts - post.MM.sun.contrasts))
```

# sessionInfo()
```{r}
sessionInfo()
```

