---
title: "Chap14_4_June29"
author: "Kazu"
date: "4/1/2020"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=TRUE,cache = FALSE)
library(rethinking)
library(lme4);library(lmerTest);library(ggplot2);library(reshape2);library(tidyverse);library(readr)
```

# 14.5.2. Example: Phylogenetic distance.
```{r}
library(rethinking)
data(Primates301)
data(Primates301_nex)
# plot it
library(ape)
### cut in a book. Use code in github
plot( ladderize(Primates301_nex) , type="fan" , font=1 , no.margin=TRUE ,
    label.offset=1 , cex=0.5 )

```

```{r}
d <- Primates301
d$name <- as.character(d$name)
dstan <- d[ complete.cases( d$group_size , d$body , d$brain ) , ]
spp_obs <- dstan$name
```

```{r}
dat_list <- list(
    N_spp = nrow(dstan),
    M = standardize(log(dstan$body)),
    B = standardize(log(dstan$brain)),
    G = standardize(log(dstan$group_size)),
    Imat = diag( nrow(dstan) )
)
```



```{r}
m14.8 <- ulam(
    alist(
        G ~ multi_normal( mu , SIGMA ),
        mu <- a + bM*M + bB*B,
        matrix[N_spp,N_spp]: SIGMA <- Imat * sigma_sq,
        a ~ normal( 0 , 1 ),
        c(bM,bB) ~ normal( 0 , 0.5 ),
        sigma_sq ~ exponential( 1 )
    ), data=dat_list , chains=4 , cores=4 )
precis( m14.8 )
```

```{r R code 14.49}
tree_trimmed <- keep.tip( Primates301_nex, spp_obs )
Rbm <- corBrownian( phy=tree_trimmed )
V <- vcv(Rbm)
Dmat <- cophenetic( tree_trimmed )
plot( Dmat , V , xlab="phylogenetic distance" , ylab="covariance" )
```

```{r R code 14.50}
# put species in right order
dat_list$V <- V[ spp_obs , spp_obs ]
# convert to correlation matrix
dat_list$R <- dat_list$V / max(V)
# Brownian motion model
m14.9 <- ulam(
    alist(
        G ~ multi_normal( mu , SIGMA ),
        mu <- a + bM*M + bB*B,
        matrix[N_spp,N_spp]: SIGMA <- R * sigma_sq,
        a ~ normal( 0 , 1 ),
        c(bM,bB) ~ normal( 0 , 0.5 ),
        sigma_sq ~ exponential( 1 )
    ), data=dat_list , chains=4 , cores=4 )
precis( m14.9 )
# heatmap (from book text)
# image(V)
# image(Dmat)
```

```{r R code 14.51}
# add scaled and reordered distance matrix
dat_list$Dmat <- Dmat[ spp_obs , spp_obs ] / max(Dmat)
m14.10 <- ulam(
    alist(
        G ~ multi_normal( mu , SIGMA ),
        mu <- a + bM*M + bB*B,
        matrix[N_spp,N_spp]: SIGMA <- cov_GPL2( Dmat , etasq , rhosq , 0.01 ),
        a ~ normal(0,1),
        c(bM,bB) ~ normal(0,0.5),
        etasq ~ exponential(1),
        rhosq ~ exponential(1)
    ), data=dat_list , chains=4 , cores=4 )
precis( m14.10 )
```

```{r R code 14.52}
 post <- extract.samples(m14.10)
plot( NULL , xlim=c(0,max(dat_list$Dmat)) , ylim=c(0,5) ,
    xlab="phylogenetic distance" , ylab="covariance" )
for ( i in 1:50 ) curve( post$etasq[i]*exp(-post$rhosq[i]*x^2) , add=TRUE , col=grau() )
```


# 14.7. Practice
# 14M5. 
* Modify the phylogenetic distance example to use group size as the outcome and brain size as
a predictor. Assuming brain size influences group size, what is your estimate of the effect? How does
phylogeny influence the estimate? 
```{r}

```


# 14H2
* which is 234 height measurements on 26 boys from an Oxford Boys Club (I think these were like youth athletic leagues?), at 9 different ages (centered and standardized) per boy. You’ll be interested in predicting height, using age, clustered by Subject (individual boy).
* Fit a model with varying intercepts and slopes (on age), clustered by Subject. Present and interpret the parameter estimates. Which varying effect contributes more variation to the heights, the intercept or the slope?
```{r}
data(Oxboys)
summary(Oxboys)
head(Oxboys)
Oxboys$Subject <- as.character(Oxboys$Subject)

# looking raw data
Oxboys %>% ggplot(aes(x=age, y=height,color=Subject)) + geom_line()
```

# 
```{r}
Oxboys.list <- list(
  Subject=as.integer(Oxboys$Subject),
  age=Oxboys$age,
  height=Oxboys$height
)
mOxboys.partialpooling <- ulam(
    alist(
        height ~ dnorm( mu , sigma ),
        mu <- a[Subject] + b[Subject]*age, 
        a[Subject] ~ dnorm(a_bar,a_sigma),
        b[Subject] ~ dnorm(0,1),  
        a_bar ~ dnorm(145,1),
        a_sigma ~ dexp(1),
        sigma ~ dexp(1)
    ) , data=Oxboys.list , chains=4 , cores=2,iter=1000,log_lik=TRUE)
# 
precis( mOxboys.partialpooling )
plot(mOxboys.partialpooling)
```

```{r}
mOxboys.multinormal <- ulam(
  alist(
    height ~ normal(mu, sigma),
    mu <- a[Subject] + bS[Subject] * age,
    c(a, bS)[Subject] ~ multi_normal(c(a_S, b_S) , Rho , sigma_S),
    a_S ~ normal(145, 1),
    b_S ~ normal(0, 1),
    sigma_S ~ exponential(1),
    sigma ~ exponential(1),
    Rho ~ lkj_corr(2)
  ),
  data = Oxboys.list,
  chains = 4,
  cores = 4,
  log_lik = T,
  iter = 5000
)
precis(mOxboys.multinormal, depth = 3, pars = c("a", "bS", "Rho", "sigma"))
```


# 14H3
* Now consider the correlation between the varying intercepts and slopes. 
* Can you explain its value? How would this estimated correlation influence your predictions about a new sample of boys?
```{r}


```


# 14H3
```{r}

```




```{r}
sessionInfo()
```

