---
title: "Summary of models"
output: html_notebook
---

# prepaå
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=TRUE,cache = FALSE)
library(rethinking)
library(lme4);library(lmerTest);library(ggplot2);library(reshape2);library(tidyverse);library(readr)
```

# Key words
* no pooling, partial pooling, complete? pooling???
* multilevel
* 



# multilevel tadpoles
```{r}
data(reedfrogs)
d <- reedfrogs
str(d)
d
# make the tank cluster variable
d$tank <- 1:nrow(d)
dat <- list(
    S = d$surv,
    N = d$density,
    tank = d$tank )

```

* varying intercepts model (= simplelst kind of varying effects) 

$$
Si \sim Bionial(N_i,p_i) \\
logit(p_i) \sim \alpha_{TANK[i]} \\
\alpha_j \sim Normal(0,1.5),\ for\ j=1..48 \\
$$

```{r}
m13.1 <- ulam(
    alist(
        S ~ dbinom( N , p ),
        logit(p) <- a[tank],
        a[tank] ~ dnorm( 0 , 1.5 )
), data=dat, chains=4, log_lik=TRUE ,cores=2) # cores=2 added by Kazu

```

* multilevel model

$$
Si \sim Bionial(N_i,p_i) \\
logit(p_i) \sim \alpha_{TANK[i]} \\
\alpha_j \sim Normal(\bar\alpha,sigma) \\
\bar\alpha \sim Normal(0,1.5) \\
\sigma \sim Exponential(1)\\
$$



```{r}
m13.2 <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[tank] ,
        a[tank] ~ dnorm( a_bar , sigma ) ,
        a_bar ~ dnorm( 0 , 1.5 ) ,
             sigma ~ dexp( 1 )
), data=dat, chains=4 , log_lik=TRUE ,cores=2) # cores=2 added by Kazu

```




# pooling BY POND (13.2.1)
* No pooling: no information is shared across ponds. It's like assuming that the variation among ponds is infinite, so nothing you learn from one pond helps you predict another.
## data fimulation

```{r}
a_bar <- 1.5; sigma <- 1.5; nponds <-60; Ni <- as.integer(rep(c(5,10,25,35),each=15))
set.seed(5005)
a_pond <- rnorm( nponds , mean=a_bar , sd=sigma )
dsim <- data.frame(pond=1:nponds, Ni=Ni ,true_a=a_pond)
dsim$Si <- rbinom(nponds, prob=logistic(dsim$true_a),size=dsim$Ni)
```
# no pooling
```{r}
dsim$p_nopool <- dsim$Si / dsim$Ni
```


* Partial pooling: 

$$
Si \sim Bionial(N_i,p_i) \\
logit(p_i) \sim \alpha_{POND[i]} \\
\alpha_j \sim Normal(\bar\alpha,sigma) \\
\bar\alpha \sim Normal(0,1.5) \\
\sigma \sim Exponential(1)\\
$$
$$\bar\alpha$$ is the average log-odds of survival in the entire population of ponds.
$$\sigma$$ the standard deviation of the distribution of log-odds of survival among ponds
$$\alpha$$, a vector of individual pond intercepts, one for each pond

# 13.2.5 compute the partial-pooling estimates

```{r}
dat <- list(Si=dsim$Si,Ni=dsim$Ni,pond=dsim$pond)
m13.3 <- ulam(
    alist(
        Si ~ dbinom( Ni , p ),
        logit(p) <- a_pond[pond],
        a_pond[pond] ~ dnorm( a_bar , sigma ),
        a_bar ~ dnorm( 0 , 1.5 ),
        sigma ~ dexp( 1 )
    ), data=dat , chains=4 )
precis( m13.3 , depth=2 )
post <- extract.samples(m13.3)
dsim$p_partpool <- apply(inv_logit(post$a_pond),2,mean)
dsim$p_true <- inv_logit(dsim$true_a)
nopool_error <- abs(dsim$p_nopool - dsim$p_true)
partpool_error <- abs(dsim$p_partpool - dsim$p_true)
plot(1:60, nopool_error,xlab="pond",ylab="absolute error",col=rangi2,pch=16)
points(1:60,partpool_error) # did not work

```

# 13.3.1 Multilevel chimpanzes

```{r}
data(chimpanzees)
d <- chimpanzees
summary(d)
d$treatment <- 1 + d$prosoc_left + 2*d$condition # see R code 11.2 for detailed explanation
dat_list<- list(
    pulled_left = d$pulled_left,
    actor = d$actor,
    block_id = d$block,
    treatment = as.integer(d$treatment) )
```



$$
L_i \sim Binomial(1,p_i) \\
logit(p_i) =  \alpha_{ACTOR[i]} + \gamma_{BLOCK[i]} + \beta_{TREATMENT[i]} \\
\beta_{j} \sim Normal(0,0.5)\ ,for\ j=1..4\\ 
$$
* there is four treatments

#### a model description under construction ##

$$
\alpha_{j} \sim Normal(\bar\alpha,\sigma_\alpha)\ ,for\ j=1..4\\ 

$$


# m13.4 
```{r R code 13.21}
library(rethinking)
data(chimpanzees)

set.seed(13)
m13.4 <- ulam(
    alist(
        pulled_left ~ dbinom( 1 , p ) ,
        logit(p) <- a[actor] + g[block_id] + b[treatment] ,
        b[treatment] ~ dnorm( 0 , 0.5 ),
        # adaptive priors
        a[actor] ~ dnorm( a_bar , sigma_a ),
        g[block_id] ~ dnorm( 0 , sigma_g ),
        # hyper-priors
        a_bar ~ dnorm( 0 , 1.5 ),
        sigma_a ~ dexp(1),
        sigma_g ~ dexp(1)
    ) , data=dat_list , chains=4 , cores=4 , log_lik=TRUE )
```

```{r}
sessionInfo()
```

